---
title: "Fairness review"
author: "Kendra Wyant"
editor_options: 
  chunk_output_type: console
format: 
  html: 
    embed-resources: true
number-sections: true
bibliography: paper\_lag.bib

---

```{r}
#| echo: false
#| message: false

library(tidyverse)
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_processed <- format_path("studydata/risk/data_processed/lag")
path_shared <- format_path("studydata/risk/data_processed/shared")
```

# Literature

## Sex 
Women are less likely to enter treatment compared to males. Sex does not predict retention, completion or outcome once in treatment. However, sex-specific predictors of treatment outcome exist [@greenfieldSubstanceAbuseTreatment2007].

Women are also more vulnerable to adverse health and social consequences of substance use disorders [@mchughSexGenderDifferences2018]. They typically show an accelerated progression from onset of use to treatment entry and develop severe symptoms over fewer years compared to males [@campbellBriefReportGender2018].   

Women are more likely to face additional barriers such as being unemployed, financially dependent, having psychiatric comorbidities, and having childcare responsibilities. Women who have children or are pregnant often report not entering treatment because they are scared to lose custody or have their baby taken away from them [@greenfieldSubstanceAbuseTreatment2007; @parlier-ahmadExploratoryStudySex2021].

Women were less likely to obtain follow-up treatment compared with men after an emergency room visit for an opioid overdose [@kilaruIncidenceTreatmentOpioid2020].

## Income 
After Affordable Care Act, those with OUD were significantly less likely to be uninsured and were less likely to report financial barriers as a reason for not receiving substance use treatment. Individuals were also more likely to receive substance use treatment and were more likely to report that insurance paid for treatment [@mckennaTreatmentUseSources2017; @johnsonReceivingAddictionTreatment2020].

Still, disparities exist. Seventeen percent of low-income adults with substance use disorders reported having no health insurance coverage in the past year. These individuals were less likely to report a past year SUD treatment visit, an outpatient care visit, an emergency department visit, or inpatient admission compared to low-income adults with continuous or discontinuous coverage [@olfsonHealthcareCoverageService2022]. 
	
## Age 
Older adults (65 years +) may have complicating factors increasing risk related to SUDs (medical comorbidities, medications and health history) [@kuerbisSubstanceAbuseOlder2014]. Additionally, other factors such as being divorced, widowed, retired, or fewer social activities could result in family or friends missing symptoms related to harmful substance use. Doctors also may attribute symptoms to other health or age related factors.

Participants over 65 were less likely to receive SUD treatment or perceive a need for treatment compared to other age groups. Additionally, they found that the 50 - 64 years age group did not differ from the 35 - 49 years age group in SUD treatment use or perceived need for treatment [@choiTreatmentUsePerceived2014]. Those in age group 26 - 49 years are more likely to receive treatment [@johnsonReceivingAddictionTreatment2020].
			
For each additional year of age, patients were 0.2% less likely to obtain follow-up treatment after an emergency room visit for an opioid overdose [@kilaruIncidenceTreatmentOpioid2020]. Additionally, middle age is associated with fewer drug related ER visits compared to younger or older people.

## Race and ethnicity

Black and Hispanic individuals are less likely to receive treatment and are more likely to experience negative consequences associated with SUDs [@pinedoCurrentReexaminationRacial2019; @guerreroDisparitiesLatinoSubstance2013; @schmidtEthnicDisparitiesClinical2007]. 

Black patients were half as likely to obtain follow-up treatment compared with non-Hispanic white patients after an emergency room visit for an opioid overdose. Hispanic patients were also less likely to obtain follow-up treatment [@kilaruIncidenceTreatmentOpioid2020].

## SUD severity
People with severe SUD are more likely to seek and receive treatment compares to those with mild or moderate SUDs [@johnsonReceivingAddictionTreatment2020]. However, those with severe SUDs may face additional barriers to completing treatment.


# Representation in our data

```{r}
#| message: false

disposition <- read_csv(file.path(path_processed, "disposition.csv"), 
                        col_types = "ccDDcccccccccc")
screen <- read_csv(file.path(path_shared, "screen.csv"), 
                   col_types = cols()) |>
  filter(subid %in% subset(disposition, analysis == "yes")$subid) |> 
  mutate(across(dsm5_1:dsm5_11, ~ recode(., "No" = 0, "Yes" = 1))) |>  
  rowwise() |>  
  mutate(dsm5_total = sum(c(dsm5_1, dsm5_2, dsm5_3, dsm5_4, dsm5_5, dsm5_6, dsm5_7, 
                              dsm5_8, dsm5_9, dsm5_10, dsm5_11))) |>  
  ungroup() 
```

### Sex (male vs. female)
```{r}
janitor::tabyl(screen$dem_2)
```


### Income (at or above poverty line vs. below poverty line)


$15,060 is federal poverty line
```{r}
screen |> 
  mutate(income = if_else(dem_7 < 15060, "below poverty line", "above poverty line")) |> 
  janitor::tabyl(income)
```

### Age (younger vs. older)

Only 3 participants 65 or older    
**8 participants 60 or older**    
21 participants 55 or older <- however lit suggests this group is different than 60+
```{r}
janitor::tabyl(screen$dem_1)
```


### Race and ethnicity

```{r}
screen <- screen |> 
  mutate(race = if_else(dem_3 == "White/Caucasian" & dem_4 == "No, I am not of Hispanic, Latino, or Spanish origin", "Non-Hispanic White", "Other"))

janitor::tabyl(screen$race)
```

### Marital status

```{r}
janitor::tabyl(screen$dem_8)
```

### Education

```{r}
janitor::tabyl(screen$dem_5)
```

### Employment

```{r}
janitor::tabyl(screen$dem_6)

janitor::tabyl(screen$dem_6_1)

```

### SUD severity
Only 7 participants have moderate AUD (6+ = severe)
```{r}
janitor::tabyl(screen$dsm5_total)
```


# References
