---
title: "SAMHSA 2019 data"
author: "Kendra Wyant"
editor_options: 
  chunk_output_type: console
format: 
  html: 
    embed-resources: true
---

```{r}
#| echo: false
#| message: false

library(tidyverse)
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_processed <- format_path("studydata/risk/data_processed/lag")
path_shared <- format_path("studydata/risk/data_processed/shared")
```

```{r}
#| message: false

load("../NSDUH_2019.Rdata")

d <- PUF2019_100920 |> 
  mutate(subid = QUESTID2,
         race = if_else(NEWRACE2 == 1, "Non-Hispanic White", "Other"),
         sex = if_else(irsex == 1, "man", "woman"),
         age = case_when(AGE2 == 1 | AGE2 == 2 ~ "12-13",
                         AGE2 == 3 |AGE2 == 4 ~ "14-15", 
                         AGE2 == 5 | AGE2 == 6 ~ "16-17",
                         AGE2 == 7 | AGE2 == 8 | AGE2 == 9 ~ "18-20",
                         AGE2 == 10 | AGE2 == 11 ~ "21-23",
                         AGE2 == 12 ~ "24-25",
                         AGE2 == 13 ~ "26-29",
                         AGE2 == 14 ~ "30-34",
                         AGE2 == 15 ~ "35-49",
                         AGE2 == 16 ~ "50-64",
                         AGE2 == 17 ~ "65+"),
         income = factor(IRPINC3, 
                         levels = 1:7, 
                         labels = c("< 10,000", "10,000 - 19,999", "20,000 - 29,999",
                                    "30,000 - 39,999", "40,000 - 49,999", "50,000 - 74,999",
                                    "75,000+") ),
         poverty = factor(POVERTY3, levels = 1:3, 
                          labels = c("poverty", "up to 2x federal poverty threshold",
                                     "more than 2x federal poverty threshold")),
         employment = factor(WRKSTATWK2,
                             levels = 1:9,
                             labels = c("full-time", "part-time", "student",
                                        "volunteer or has job but did not work in past week",
                                        "unemployed/looking for work", "disabled",
                                        "keeping house full-time", "retired",
                                        "not working for other reason")),
         marital = factor(irmarit, 
                          levels = 1:4,
                          labels = c("married", "widowed", "divorced or separated",
                                     "never been married")),
         education = factor(IREDUHIGHST2,
                            levels = 1:11,
                            labels = c("fifth grade or less",
                                       "sixth grade",
                                       "seventh grade",
                                       "eighth grade",
                                       "ninth grade",
                                       "tenth grade",
                                       "eleventh or twelfth grade (no diploma)",
                                       "high school diploma/GED",
                                       "some college",
                                       "Associate's degree",
                                       "college graduate or higher")),
         aud_past_year = if_else(depndalc == 1, "yes", "no"),
         # aud_sev_past_year = factor(IRPYSEV5ALC, levels = c(1,2,3,9),
         # labels = c("mild", "moderate", "severe", "none")),
         aud_tx_past_year = if_else(txyraldgb == 1 | txyraldgb == 3 |
                                      txyraldgb == 4 | txyraldgb == 6 |
                                      txyraldgb == 11, "yes", "no"))
```

## AUD prevalence

3% of individuals in data set met criteria for AUD "DEPENDENCE"
```{r}
janitor::tabyl(d$aud_past_year)
```

Going to filter out those without AUD
```{r}
d <- d |> 
  filter(aud_past_year == "yes")
```

## Obtained tx

### Any alcohol use treatment reported

```{r}
janitor::tabyl(d$aud_tx_past_year)
```

### Race/ethnicity

Non-Hispanic Whites receive treatment more than other races as an aggregate other category
```{r}
janitor::tabyl(d, race, aud_tx_past_year) 

janitor::tabyl(d, race, aud_tx_past_year) |> 
  janitor::adorn_percentages("row")
```

### Sex (male vs. female)   

Gender access to tx appears equal
```{r} 
janitor::tabyl(d, sex, aud_tx_past_year) 

janitor::tabyl(d, sex, aud_tx_past_year) |> 
  janitor::adorn_percentages("row")
```

### Income 

personal income brackets 
```{r}
janitor::tabyl(d, income, aud_tx_past_year) 

janitor::tabyl(d, income, aud_tx_past_year) |> 
  janitor::adorn_percentages("row")
```

poverty determined by family income    

- this variable (POVERTY3) was created for each individual based on their poverty threshold (determined by their age, family size, the number of children in the household, and total family income). Since NSDUH only collects categorical values of family income, for each level of income, the midpoint of each range was assigned as the effective family income.
```{r}
janitor::tabyl(d, poverty, aud_tx_past_year) 

janitor::tabyl(d, poverty, aud_tx_past_year) |> 
  janitor::adorn_percentages("row")
```

### Age 

Most granular age categories
```{r}
janitor::tabyl(d, age, aud_tx_past_year) 

janitor::tabyl(d, age, aud_tx_past_year) |> 
  janitor::adorn_percentages("row") |> 
  arrange(age)
```

### Marital status

```{r}
janitor::tabyl(d, marital, aud_tx_past_year) 

janitor::tabyl(d, marital, aud_tx_past_year) |> 
  janitor::adorn_percentages("row")
```

### Education

```{r}
janitor::tabyl(d, education, aud_tx_past_year) 

janitor::tabyl(d, education, aud_tx_past_year) |> 
  janitor::adorn_percentages("row")
```

### Employment

```{r}
janitor::tabyl(d, employment, aud_tx_past_year) 

janitor::tabyl(d, employment, aud_tx_past_year) |> 
  janitor::adorn_percentages("row")
```

