---
title: "SAMHSA 2018 data"
author: "Kendra Wyant"
editor_options: 
  chunk_output_type: console
format: 
  html: 
    embed-resources: true
---

```{r}
#| echo: false
#| message: false

library(tidyverse)
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_processed <- format_path("studydata/risk/data_processed/lag")
path_shared <- format_path("studydata/risk/data_processed/shared")
```

```{r}
#| message: false

load("../NSDUH_2018.Rdata")

d <- NSDUH_2018 |> 
  mutate(subid = QUESTID2,
         race = if_else(NEWRACE2 == 1, "Non-Hispanic White", "Other"),
         sex = if_else(irsex == 1, "man", "woman"),
         age = case_when(AGE3 == 1 ~ "12-13",
                         AGE3 == 2 ~ "14-15", 
                         AGE3 == 3 ~ "16-17",
                         AGE3 == 4 ~ "18-20",
                         AGE3 == 5 ~ "21-23",
                         AGE3 == 6 ~ "24-25",
                         AGE3 == 7 ~ "26-29",
                         AGE3 == 8 ~ "30-34",
                         AGE3 == 9 ~ "35-49",
                         AGE3 == 10 ~ "50-64",
                         AGE3 == 11 ~ "65+"),
         income = factor(IRPINC3, 
                         levels = 1:7, 
                         labels = c("< 10,000", "10,000 - 19,999", "20,000 - 29,999",
                                    "30,000 - 39,999", "40,000 - 49,999", "50,000 - 74,999",
                                    "75,000+") ),
         poverty = factor(POVERTY3, levels = 1:3, 
                          labels = c("poverty", "up to 2x federal poverty threshold",
                                     "more than 2x federal poverty threshold")),
         employment = factor(WRKSTATWK2,
                             levels = 1:9,
                             labels = c("full-time", "part-time", "student",
                                        "volunteer or has job but did not work in past week",
                                        "unemployed/looking for work", "disabled",
                                        "keeping house full-time", "retired",
                                        "not working for other reason")),
         marital = factor(irmarit, 
                          levels = 1:4,
                          labels = c("married", "widowed", "divorced or separated",
                                     "never been married")),
         education = factor(IREDUHIGHST2,
                            levels = 1:11,
                            labels = c("fifth grade or less",
                                       "sixth grade",
                                       "seventh grade",
                                       "eighth grade",
                                       "ninth grade",
                                       "tenth grade",
                                       "eleventh or twelfth grade (no diploma)",
                                       "high school diploma/GED",
                                       "some college",
                                       "Associate's degree",
                                       "college graduate or higher")),
         aud_past_year = if_else(IRPYUD5ALC == 1, "yes", "no"),
         aud_sev_past_year = factor(IRPYSEV5ALC, levels = c(1,2,3,9),
         labels = c("mild", "moderate", "severe", "none")),
         aud_tx_clinician_past_year = if_else(irsutinhosp == 1 | irsutinrhab == 1 |
                                               irsutinrsmh == 1 | irsutinopl == 1 | irsutoutrhab == 1 |
                                                irsutoutmhcr == 1 | irsutoutthrp == 1 | irsutoutdoc == 1 |
                                                irsutouthosp == 1 | irsutoutschl == 1 | irsutoutopl == 1 |
                                                irsutrxalc == 1 | irsutphvid == 1 | irsutdetox == 1,
                                    "yes", "no"),
         aud_tx_clinician_carceral_past_year = if_else(irsutinhosp == 1 | irsutinrhab == 1 |
                                               irsutinrsmh == 1 | irsutinopl == 1 | irsutoutrhab == 1 |
                                                irsutoutmhcr == 1 | irsutoutthrp == 1 | irsutoutdoc == 1 |
                                                irsutouthosp == 1 | irsutoutschl == 1 | irsutoutopl == 1 |
                                                irsutrxalc == 1 | irsutphvid == 1 | irsutdetox == 1 |
                                               sutrtalcpy == 1 | sutrtdralpy == 1,
                                    "yes", "no"),
         aud_tx_any_past_year = if_else(irsutinhosp == 1 | irsutinrhab == 1 | irsutinrsmh == 1 | 
                                          irsutinopl == 1 | irsutoutrhab == 1 | irsutoutmhcr == 1 |
                                          irsutoutthrp == 1 | irsutoutdoc == 1 | irsutouthosp == 1 |
                                          irsutoutschl == 1 | irsutoutopl == 1 | irsutrxalc == 1 |
                                          irsutphvid == 1 | irsutdetox == 1  | irsutpeerspc == 1 |
                                          irsutsuppgrp == 1,
                                    "yes", "no"))
```

## AUD prevalence

10% of individuals in data set met criteria for AUD
```{r}
janitor::tabyl(d$aud_past_year)
```

Going to filter out those without AUD
```{r}
d <- d |> 
  filter(aud_past_year == "yes")
```

## Obtained tx

### Clinican delivered
Defined as treatment for alcohol or alcohol and drugs combined in the following settings:    
- outpatient
- inpatient 
- professional tx over phone or through video
- medication tx
- detox

```{r}
janitor::tabyl(d$aud_tx_clinician_past_year)
```


### Clinician delivered (including carceral settings)

Note: can't be entirely sure about the precision of this distinction

```{r}
janitor::tabyl(d$aud_tx_clinician_carceral_past_year)
```


### Any treatment

Includes the addition of support groups and peer support specialist or recovery coaches
```{r}
janitor::tabyl(d$aud_tx_any_past_year)
```

## Clinician delivered txs (not including carceral settings)
### Race and ethnicity

Non-Hispanic Whites receive treatment more than other races as an aggregate other category
```{r}
janitor::tabyl(d, race, aud_tx_clinician_past_year) 

janitor::tabyl(d, race, aud_tx_clinician_past_year) |> 
  janitor::adorn_percentages("row")
```

### Sex (male vs. female)   

Gender access to tx appears equal
```{r} 
janitor::tabyl(d, sex, aud_tx_clinician_past_year) 

janitor::tabyl(d, sex, aud_tx_clinician_past_year) |> 
  janitor::adorn_percentages("row")
```

### Income 

personal income brackets 
```{r}
janitor::tabyl(d, income, aud_tx_clinician_past_year) 

janitor::tabyl(d, income, aud_tx_clinician_past_year) |> 
  janitor::adorn_percentages("row")
```

poverty determined by family income    

- this variable (POVERTY3) was created for each individual based on their poverty threshold (determined by their age, family size, the number of children in the household, and total family income). Since NSDUH only collects categorical values of family income, for each level of income, the midpoint of each range was assigned as the effective family income.
```{r}
janitor::tabyl(d, poverty, aud_tx_clinician_past_year) 

janitor::tabyl(d, poverty, aud_tx_clinician_past_year) |> 
  janitor::adorn_percentages("row")
```

### Age 

Most granular age categories
```{r}
janitor::tabyl(d, age, aud_tx_clinician_past_year) 

janitor::tabyl(d, age, aud_tx_clinician_past_year) |> 
  janitor::adorn_percentages("row") |> 
  arrange(age)
```

### Marital status

```{r}
janitor::tabyl(d, marital, aud_tx_clinician_past_year) 

janitor::tabyl(d, marital, aud_tx_clinician_past_year) |> 
  janitor::adorn_percentages("row")
```

### Education

```{r}
janitor::tabyl(d, education, aud_tx_clinician_past_year) 

janitor::tabyl(d, education, aud_tx_clinician_past_year) |> 
  janitor::adorn_percentages("row")
```

### Employment

```{r}
janitor::tabyl(d, employment, aud_tx_clinician_past_year) 

janitor::tabyl(d, employment, aud_tx_clinician_past_year) |> 
  janitor::adorn_percentages("row")
```

### SUD severity

```{r}
janitor::tabyl(d, aud_sev_past_year, aud_tx_clinician_past_year) 

janitor::tabyl(d, aud_sev_past_year, aud_tx_clinician_past_year) |> 
  janitor::adorn_percentages("row")
```



## Clinician delivered txs (including carceral settings)
### Race and ethnicity

Non-Hispanic Whites receive treatment more than other races as an aggregate other category
```{r}
janitor::tabyl(d, race, aud_tx_clinician_carceral_past_year) 

janitor::tabyl(d, race, aud_tx_clinician_carceral_past_year) |> 
  janitor::adorn_percentages("row")
```

### Sex (male vs. female)   

Gender access to tx appears equal
```{r} 
janitor::tabyl(d, sex, aud_tx_clinician_carceral_past_year) 

janitor::tabyl(d, sex, aud_tx_clinician_carceral_past_year) |> 
  janitor::adorn_percentages("row")
```

### Income 

personal income brackets 
```{r}
janitor::tabyl(d, income, aud_tx_clinician_carceral_past_year) 

janitor::tabyl(d, income, aud_tx_clinician_carceral_past_year) |> 
  janitor::adorn_percentages("row")
```

poverty determined by family income    

- this variable (POVERTY3) was created for each individual based on their poverty threshold (determined by their age, family size, the number of children in the household, and total family income). Since NSDUH only collects categorical values of family income, for each level of income, the midpoint of each range was assigned as the effective family income.
```{r}
janitor::tabyl(d, poverty, aud_tx_clinician_carceral_past_year) 

janitor::tabyl(d, poverty, aud_tx_clinician_carceral_past_year) |> 
  janitor::adorn_percentages("row")
```

### Age 

Most granular age categories
```{r}
janitor::tabyl(d, age, aud_tx_clinician_carceral_past_year) 

janitor::tabyl(d, age, aud_tx_clinician_carceral_past_year) |> 
  janitor::adorn_percentages("row") |> 
  arrange(age)
```

### Marital status

```{r}
janitor::tabyl(d, marital, aud_tx_clinician_carceral_past_year) 

janitor::tabyl(d, marital, aud_tx_clinician_carceral_past_year) |> 
  janitor::adorn_percentages("row") 
```

### Education

```{r}
janitor::tabyl(d, education, aud_tx_clinician_carceral_past_year) 

janitor::tabyl(d, education, aud_tx_clinician_carceral_past_year) |> 
  janitor::adorn_percentages("row")
```

### Employment

```{r}
janitor::tabyl(d, employment, aud_tx_clinician_carceral_past_year) 

janitor::tabyl(d, employment, aud_tx_clinician_carceral_past_year) |> 
  janitor::adorn_percentages("row")
```

### SUD severity

```{r}
janitor::tabyl(d, aud_sev_past_year, aud_tx_clinician_carceral_past_year) 

janitor::tabyl(d, aud_sev_past_year, aud_tx_clinician_carceral_past_year) |> 
  janitor::adorn_percentages("row")
```

## Any treatment
### Race and ethnicity

Non-Hispanic Whites receive treatment more than other races as an aggregate other category
```{r}
janitor::tabyl(d, race, aud_tx_any_past_year) 

janitor::tabyl(d, race, aud_tx_any_past_year) |> 
  janitor::adorn_percentages("row")
```

### Sex (male vs. female)   

Gender access to tx appears equal
```{r} 
janitor::tabyl(d, sex, aud_tx_any_past_year) 

janitor::tabyl(d, sex, aud_tx_any_past_year) |> 
  janitor::adorn_percentages("row")
```

### Income 

personal income brackets 
```{r}
janitor::tabyl(d, income, aud_tx_any_past_year) 

janitor::tabyl(d, income, aud_tx_any_past_year) |> 
  janitor::adorn_percentages("row")
```

poverty determined by family income    

- this variable (POVERTY3) was created for each individual based on their poverty threshold (determined by their age, family size, the number of children in the household, and total family income). Since NSDUH only collects categorical values of family income, for each level of income, the midpoint of each range was assigned as the effective family income.
```{r}
janitor::tabyl(d, poverty, aud_tx_any_past_year) 

janitor::tabyl(d, poverty, aud_tx_any_past_year) |> 
  janitor::adorn_percentages("row")
```

### Age 

Most granular age categories
```{r}
janitor::tabyl(d, age, aud_tx_any_past_year) 

janitor::tabyl(d, age, aud_tx_any_past_year) |> 
  janitor::adorn_percentages("row") |> 
  arrange(age)
```

### Marital status

```{r}
janitor::tabyl(d, marital, aud_tx_any_past_year) 

janitor::tabyl(d, marital, aud_tx_any_past_year) |> 
  janitor::adorn_percentages("row") 
```

### Education

```{r}
janitor::tabyl(d, education, aud_tx_any_past_year) 

janitor::tabyl(d, education, aud_tx_any_past_year) |> 
  janitor::adorn_percentages("row")
```

### Employment

```{r}
janitor::tabyl(d, employment, aud_tx_any_past_year) 

janitor::tabyl(d, employment, aud_tx_any_past_year) |> 
  janitor::adorn_percentages("row")
```

### SUD severity

```{r}
janitor::tabyl(d, aud_sev_past_year, aud_tx_any_past_year) 

janitor::tabyl(d, aud_sev_past_year, aud_tx_any_past_year) |> 
  janitor::adorn_percentages("row")
```

