---
title: "Supplemental Material"
author: "Kendra Wyant, Gaylen E. Fronk, Jiachen Yu, Claire E. Punturieri, and John J. Curtin"
date: last-modified
number-sections: true
format: 
  html: 
    embed-resources: true
    toc: true 
    toc_depth: 5
execute:
  echo: false
editor_options: 
  chunk_output_type: console
---

This file contains the supplemental materials for *Forecasting Alcohol Lapse Risk up to Two Weeks in Advance using Time-lagged Machine Learning Models*. It includes all supplemental figures and tables. Additional materials are made available on our study's OSF page ([https://osf.io/xta67/](https://osf.io/xta67/)).   

-----

## Supplemental Figures

### Figure S1: Raw Predicted Probabilities

```{r}
#| code-fold: true

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true"))
suppressPackageStartupMessages(library(tidyposterior))
suppressPackageStartupMessages(library(kableExtra))
suppressWarnings(library(ggrepel))
library(patchwork)


theme_set(theme_classic())

options(knitr.kable.NA = '')

path_models_lag <- format_path(str_c("risk/models/lag"))
path_processed <- format_path(str_c("risk/data_processed/lag"))

pp_tidy <- read_csv(here::here(path_models_lag, "posteriors.csv"), 
                                 show_col_types = FALSE) 

ci <- read_csv(here::here(path_models_lag, "pp_perf_tibble.csv"), 
                                 show_col_types = FALSE) |> 
  mutate(model = factor(model, levels = c("0 lag", "24 lag", "72 lag", "168 lag", "336 lag"),
                        labels = c("No lag", "1-day", "3-day", "1-week", "2-week")))

pp_dem <- read_csv(here::here(path_models_lag, "pp_dem_all.csv"), 
                                 show_col_types = FALSE)

pp_dem_contrast <- read_csv(here::here(path_models_lag, "pp_dem_contrast_all.csv"), 
                                 show_col_types = FALSE) |> 
  mutate(lag = factor(lag, levels = c("0", "24", "72", "168", "336"),
                        labels = c("No lag", "1-day", "3-day", "1-week", "2-week")))

labels <- read_csv(here::here(path_processed, "labels_1day_0lag.csv"),
                   show_col_types = FALSE)

dem_tbl <- read_csv(here::here(path_processed, "dem_tibble.csv"),
                   show_col_types = FALSE)

labels_dem <- labels |> 
  left_join(dem_tbl, by = "subid")


preds_0 <- read_rds(here::here(path_models_lag, 
                               "final_preds_kfold_1_x_5_1day_0_v3_strat_lh_final.rds")) 
preds_24 <- read_rds(here::here(path_models_lag, 
                               "final_preds_kfold_1_x_5_1day_24_v3_strat_lh_final.rds")) 
preds_72 <- read_rds(here::here(path_models_lag, 
                               "final_preds_kfold_1_x_5_1day_72_v3_strat_lh_final.rds")) 
preds_168 <- read_rds(here::here(path_models_lag, 
                               "final_preds_kfold_1_x_5_1day_168_v3_strat_lh_final.rds")) 
preds_336 <- read_rds(here::here(path_models_lag, 
                               "final_preds_kfold_1_x_5_1day_336_v3_strat_lh_final.rds")) 

preds_all <- preds_0 |> 
  mutate(model = "No lag") |> 
  bind_rows(preds_24 |> 
              mutate(model = "1-day")) |> 
  bind_rows(preds_72 |> 
              mutate(model = "3-day")) |> 
  bind_rows(preds_168 |> 
              mutate(model = "1-week")) |> 
  bind_rows(preds_336 |> 
              mutate(model = "2-week")) |> 
  mutate(model = factor(model, levels = c("No lag", "1-day", "3-day", "1-week", "2-week")))

preds_hist <- preds_all |> 
  ggplot(aes(x = prob_raw)) +
  geom_histogram(color = "black", fill = "light grey", binwidth = .10,
                 boundary = 0) +
  facet_wrap(~model) +
  labs(x = "Raw probabilities from final model") +
  scale_x_continuous(breaks = seq(0, 1, by = 0.2))
```

Our models produced variation in predicted probabilities, with probabilities spanning nearly the entire 0 - 1 range. The 2-week lagged model had fewer high raw probabilities (all probabilities were < .80).  Histograms of the raw predicted probabilities for the best configuration for all 5 models are presented below. 

```{r}
#| label: fig-s1
#| fig-cap: "Histograms of raw probabilities of our best model configurations for each classification model (no lag, 1-day, 3-day, 1-week and 2-week lags)."
#| fig-width: 8
#| fig-height: 5

preds_hist
```

### Figure S2: Calibration Plots for all Models

The raw probabilities produced by our final models were not well calibrated. Beta and logistic calibration methods improved probability accuracy. Below we present the calibration plots of our raw predicted lapse probabilities for each model. In supplemental table S2 we report the brier scores for raw and calibrated probabilities for each model.

```{r}
#| code-fold: true

bin_width = 0.10

preds_all_raw <- preds_0 |> 
  mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "Lapse", 1, 0)) |> 
  mutate(model = "No lag") |> 
  bind_rows(preds_24 |> 
              mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)), 
                     lapse = if_else(label == "Lapse", 1, 0)) |> 
              mutate(model = "1-day")) |> 
  bind_rows(preds_72 |> 
              mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)), 
                     lapse = if_else(label == "Lapse", 1, 0)) |> 
              mutate(model = "3-day")) |> 
  bind_rows(preds_168 |> 
              mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)), 
                     lapse = if_else(label == "Lapse", 1, 0)) |> 
              mutate(model = "1-week")) |> 
  bind_rows(preds_336 |> 
              mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)), 
                     lapse = if_else(label == "Lapse", 1, 0)) |> 
              mutate(model = "2-week")) |> 
  mutate(model = factor(model, levels = c("No lag", "1-day", "3-day", "1-week", "2-week")),
         prob = "Raw (uncalibrated)")

preds_all_logi <- preds_0 |> 
  mutate(bins = cut(prob_logi, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "Lapse", 1, 0)) |> 
  mutate(model = "No lag") |> 
  bind_rows(preds_24 |> 
              mutate(bins = cut(prob_logi, breaks = seq(0, 1, bin_width)), 
                     lapse = if_else(label == "Lapse", 1, 0)) |> 
              mutate(model = "1-day")) |> 
  bind_rows(preds_72 |> 
              mutate(bins = cut(prob_logi, breaks = seq(0, 1, bin_width)), 
                     lapse = if_else(label == "Lapse", 1, 0)) |> 
              mutate(model = "3-day")) |> 
  bind_rows(preds_168 |> 
              mutate(bins = cut(prob_logi, breaks = seq(0, 1, bin_width)), 
                     lapse = if_else(label == "Lapse", 1, 0)) |> 
              mutate(model = "1-week")) |> 
  bind_rows(preds_336 |> 
              mutate(bins = cut(prob_logi, breaks = seq(0, 1, bin_width)), 
                     lapse = if_else(label == "Lapse", 1, 0)) |> 
              mutate(model = "2-week")) |> 
  mutate(model = factor(model, levels = c("No lag", "1-day", "3-day", "1-week", "2-week")),
         prob = "Platt calibration")

preds_all <- preds_all_raw |> 
  bind_rows(preds_all_logi) 

mini_hist <- preds_all_raw |>
  count(bins, model) |> 
  bind_rows(tibble(bins = "(0.9,1]",
                   model = "2-week",
                   n = 0)) |> 
  group_by(model) |> 
  mutate(
    bins = seq(1, 10, 1),
    x = bin_width / 2 + bin_width * (bins - 1),       
    y = n / max(n) * 0.07
  ) |> 
  mutate(model = factor(model, levels = c("No lag", "1-day", "3-day",
                                          "1-week", "2-week"))) 


plot_cal <- preds_all |> 
  mutate(prob = factor(prob, levels = c("Raw (uncalibrated)", 
                                        "Platt calibration"))) |> 
  group_by(bins, model, prob)  |> 
  summarize(mean_lapse = mean(lapse), .groups = "drop") |> 
  mutate(
    bins = as.numeric(bins),
    midpoints = bin_width/2 + bin_width * (bins - 1)
  )  |> 
  ggplot(aes(x = midpoints, y = mean_lapse, group = prob,
             color = prob, linetype = prob)) +
  geom_abline(slope = 1, intercept = 0, linetype = "longdash", color = "gray80") +
  geom_line(linewidth = .7) +
  geom_rect(data = mini_hist, inherit.aes = FALSE,
            aes(xmin = x - bin_width / 2,
                xmax = x + bin_width / 2,
                ymin = -y,
                ymax = 0),
            alpha = .4) +
  facet_wrap(~ model, ncol = 2) +
  labs(
    x = "Predicted Lapse Probability (Bin Midpoint)",
    y = "Observed Lapse Probability",
    color = NULL,
    linetype = NULL
  ) +
  scale_x_continuous(breaks = seq(0, 1, bin_width), limits = c(0, 1)) +
  scale_y_continuous(limits = c(-0.08, 1), breaks = seq(0,1, bin_width),
                     expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  scale_color_manual(values = c("black", "#cb6bce")) +
  scale_linetype_manual(values = c("solid", "twodash")) +
  theme_classic(base_size = 11) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 11),
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 10),
    axis.text = element_text(size = 10),
    axis.title = element_text(face = "bold")
  )
```

```{r}
#| label: fig-s2
#| fig-cap: "Calibration plots of raw and calibrated predicted lapse probabilities (beta, isotonic, and logistic) across models (no lag, 1-day, 3-day, 1-week, and 2-week lag). Predicted probabilities (x-axis) are binned into deciles. Observed lapse probability (y-axis) represents the proportion of actual lapses observed in each bin. The dashed diagonal represents perfect calibration. Points below the line indicate overestimation and points above the line indicate underestimation. Raw probabilities are depicted as the solid black curves. Platt calibrated probabilites are depicted as the pink dashed curves. The grey histogram along the bottom of the plot represents the proportion of raw probabilities in each bin."
#| fig-width: 8
#| fig-height: 10

plot_cal
```


### Figure S3: Global Shapley Plots for all 5 Models

The top globally important feature category (i.e., highest mean |Shapley value|) for all models was past use. Future efficacy was a strong predictor for more immediate model predictions (i.e., no lag and 1-day lag), but as lag time increased importance attenuated. On the other hand, as lag time increased past/future risky situations increased in importance. Craving was relatively consistently important, in magnitude, across all models.

```{r}
#| code-fold: true

shap_grp_0 <- read_rds(here::here(path_models_lag, 
                      "final_shapsgrp_kfold_1_x_5_1day_0_v3_strat_lh_final.rds")) |> 
   filter(str_detect(variable_grp, "EMA")) |> 
   mutate(variable_grp = str_remove(variable_grp, "\\(EMA item\\)")) 


shap_grp_24 <- read_rds(here::here(path_models_lag, 
                      "final_shapsgrp_kfold_1_x_5_1day_24_v3_strat_lh_final.rds")) |> 
   filter(str_detect(variable_grp, "EMA")) |> 
   mutate(variable_grp = str_remove(variable_grp, "\\(EMA item\\)")) 


shap_grp_72 <- read_rds(here::here(path_models_lag, 
                      "final_shapsgrp_kfold_1_x_5_1day_72_v3_strat_lh_final.rds")) |> 
   filter(str_detect(variable_grp, "EMA")) |> 
   mutate(variable_grp = str_remove(variable_grp, "\\(EMA item\\)")) 


shap_grp_168 <- read_rds(here::here(path_models_lag, 
                      "final_shapsgrp_kfold_1_x_5_1day_168_v3_strat_lh_final.rds")) |> 
   filter(str_detect(variable_grp, "EMA")) |> 
   mutate(variable_grp = str_remove(variable_grp, "\\(EMA item\\)")) 

shap_grp_336 <- read_rds(here::here(path_models_lag, 
                      "final_shapsgrp_kfold_1_x_5_1day_336_v3_strat_lh_final.rds")) |> 
   filter(str_detect(variable_grp, "EMA")) |> 
  mutate(variable_grp = str_remove(variable_grp, "\\(EMA item\\)"))  

shap_global <- shap_grp_0 |>
  mutate(model = 0) |>
  bind_rows(shap_grp_24 |>
              mutate(model = 24)) |>
  bind_rows(shap_grp_72 |>
              mutate(model = 72)) |>
  bind_rows(shap_grp_168 |>
              mutate(model = 168)) |>
  bind_rows(shap_grp_336 |>
              mutate(model = 336)) |>
  group_by(variable_grp, model) |>
  summarize(mean_value = mean(abs(value)), .groups = "drop") |> 
  mutate(variable_grp = str_remove(variable_grp, "past/future "),
         variable_grp = if_else(variable_grp == "past pleasant event ",
                                "pleasant event", variable_grp))

shap_levels <- shap_global |>
  mutate(variable_grp = reorder(variable_grp, mean_value, sum)) |>
  pull(variable_grp) |>
  levels()

label_data <- shap_global |>
  filter(model == 336)

fig_s3 <- shap_global |>
  mutate(variable_grp = factor(variable_grp, levels = shap_levels)) |> 
  ggplot() +
  geom_line(aes(x = model, y = mean_value, color = variable_grp)) +
  labs(y = "Mean(|Shapley Value|)",
       x = NULL,
       fill = NULL) +
  geom_text_repel(
    data = label_data,
    aes(x = model, y = mean_value, color = variable_grp, label = variable_grp),
    fontface = "bold",
    size = 4,
    direction = "y",
    hjust = 0,
    nudge_x = 5,
    segment.size = .7,
    segment.alpha = .5,
    segment.linetype = "dotted",
    box.padding = 0,
    segment.curvature = -0.1,
    segment.ncp = 3,
    segment.angle = 20
  ) +
  scale_x_continuous(breaks = c(0, 24, 72, 168, 336),
                     labels = c("No lag", "1-day", "3-day", "1-week", "2-week"),
                     limits = c(0, 425)) +
  scale_color_manual(values = c("maroon", "Navy", "#3cb44b", 
                                "magenta", "cyan", "#9A6324", "purple")) +
  theme(axis.text=element_text(size=9.5),
        legend.key.size = unit(0.25, "cm"),
        legend.position = "none")
```

```{r}
#| label: fig-s3
#| fig-cap: "Global importance (mean |Shapley value|) for feature categories by lag time."
#| fig-width: 8
#| fig-height: 5

fig_s3
```



### Figure S4: Global Shapley Plots for Baseline and 2-Week Lag Model by Demographic Group

We saw similar patterns in global feature importance across different demographic subgroups (race/ethnicity, income, sex at birth). Plots of global feature importance by demographic subgroup for the no lag and 2-week lagged models are presented below. 

```{r}
#| code-fold: true

feat_0 <- read_csv(here::here(path_processed, 
                            "features_0lag_v3.csv"),
                 col_types = cols()) |> 
  select(label_num, subid, dttm_label, lapse, demo_age, demo_sex, 
         demo_income, demo_race) |> 
  arrange(label_num) 


feat_shaps_0 <- feat_0 |> 
  rename(id_obs = label_num) |>
  right_join(shap_grp_0, by = c("id_obs")) 

baseline_sex <- feat_shaps_0 |> 
  mutate(group = if_else(demo_sex == "Male", 
                        "advantaged group (non-Hispanic White, male, above poverty line)",
                         "disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                         ),
         group = factor(group, levels = c("disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                                          "advantaged group (non-Hispanic White, male, above poverty line)")),
         model = "No Lag") |> 
  group_by(model, group, variable_grp) |> 
  summarize(mean_value = mean(abs(value)), .groups = "drop") |> 
  arrange(mean_value) |> 
  mutate(variable_grp = factor(variable_grp),
         variable_grp = fct_inorder(variable_grp)) |> 
  ggplot() +
  geom_bar(aes(x = variable_grp, y = mean_value, group = group, 
                 fill = group),  
             stat = "identity", position = "dodge") +
  facet_wrap(~model, strip.position = "left") +
  labs(title = "Sex at Birth",
       y = NULL,
       x = NULL,
       color = NULL) +
  theme(axis.text=element_text(size=9.5),
        legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        strip.placement = "outside") +
  scale_fill_manual(values = c("#cb6bce", "#240e31")) +
  coord_flip()

baseline_race <- feat_shaps_0 |> 
  mutate(group = if_else(demo_race == "White/Caucasian", 
                         "advantaged group (non-Hispanic White, male, above poverty line)",
                         "disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                         ),
         group = factor(group, levels = c("disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                                          "advantaged group (non-Hispanic White, male, above poverty line)"))) |> 
  group_by(group, variable_grp) |> 
  summarize(mean_value = mean(abs(value)), .groups = "drop") |> 
  arrange(mean_value) |> 
  mutate(variable_grp = factor(variable_grp),
         variable_grp = fct_inorder(variable_grp)) |> 
  ggplot() +
  geom_bar(aes(x = variable_grp, y = mean_value, group = group, 
                 fill = group),  
             stat = "identity", position = "dodge") +
  labs(title = "Race/Ethnicity",
       y = "Mean(|Shapley Value|)",
       x = NULL,
       color = NULL) +
  theme(axis.text=element_text(size=9.5),
        axis.text.y = element_blank(),
        legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
  scale_fill_manual(values = c("#cb6bce", "#240e31")) +
  coord_flip()


baseline_income <- feat_shaps_0 |> 
  mutate(group = if_else(demo_income >= 1560, 
                         "advantaged group (non-Hispanic White, male, above poverty line)",
                         "disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                         ),
         group = factor(group, levels = c("disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                                          "advantaged group (non-Hispanic White, male, above poverty line)"))) |> 
  group_by(group, variable_grp) |> 
  summarize(mean_value = mean(abs(value)), .groups = "drop") |> 
  arrange(mean_value) |> 
  mutate(variable_grp = factor(variable_grp),
         variable_grp = fct_inorder(variable_grp)) |> 
  ggplot() +
  geom_bar(aes(x = variable_grp, y = mean_value, group = group, 
                 fill = group),  
             stat = "identity", position = "dodge") +
  labs(title = "Income",
       y = NULL,
       x = NULL,
       color = NULL) +
  theme(axis.text=element_text(size=9.5),
        axis.text.y = element_blank(),
        legend.position = "none",
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
  scale_fill_manual(values = c("#cb6bce", "#240e31")) +
  coord_flip()


no_lag <- baseline_sex + baseline_race + baseline_income 


feat_336 <- read_csv(here::here(path_processed, 
                            "features_336lag_v3.csv"),
                 col_types = cols()) |> 
  select(label_num, subid, dttm_label, lapse, demo_age, demo_sex, 
         demo_income, demo_race) |> 
  arrange(label_num) 


feat_shaps_336 <- feat_336 |> 
  rename(id_obs = label_num) |>
  right_join(shap_grp_336, by = c("id_obs")) 

twoweek_sex <- feat_shaps_336 |> 
  mutate(group = if_else(demo_sex == "Male", 
                         "advantaged group (non-Hispanic White, male, above poverty line)",
                         "disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                         ),
         group = factor(group, levels = c("disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                                          "advantaged group (non-Hispanic White, male, above poverty line)")),
         model = "2-Week Lag") |> 
  group_by(model, group, variable_grp) |> 
  summarize(mean_value = mean(abs(value)), .groups = "drop") |> 
  arrange(mean_value) |> 
  mutate(variable_grp = factor(variable_grp),
         variable_grp = fct_inorder(variable_grp)) |> 
  ggplot() +
  geom_bar(aes(x = variable_grp, y = mean_value, group = group, 
                 fill = group),  
             stat = "identity", position = "dodge") +
  facet_wrap(~model, strip.position = "left") +
  labs(title = "Sex at Birth",
       y = NULL,
       x = NULL,
       color = NULL) +
  theme(axis.text=element_text(size=9.5),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
        strip.placement = "outside") +
  scale_fill_manual(values = c("#cb6bce", "#240e31")) +
  coord_flip()

twoweek_race <- feat_shaps_336 |> 
  mutate(group = if_else(demo_race == "White/Caucasian", 
                         "advantaged group (non-Hispanic White, male, above poverty line)",
                         "disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                         ),
         group = factor(group, levels = c("disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                                          "advantaged group (non-Hispanic White, male, above poverty line)"))) |> 
  group_by(group, variable_grp) |> 
  summarize(mean_value = mean(abs(value)), .groups = "drop") |> 
  arrange(mean_value) |> 
  mutate(variable_grp = factor(variable_grp),
         variable_grp = fct_inorder(variable_grp)) |> 
  ggplot() +
  geom_bar(aes(x = variable_grp, y = mean_value, group = group, 
                 fill = group),  
             stat = "identity", position = "dodge") +
  labs(title = "Race/Ethnicity",
       y = "Mean(|Shapley Value|)",
       x = NULL,
       color = NULL) +
  theme(axis.text=element_text(size=9.5),
        axis.text.y = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
  scale_fill_manual(values = c("#cb6bce", "#240e31")) +
  coord_flip()

twoweek_income <- feat_shaps_336 |> 
  mutate(group = if_else(demo_income >= 1560, 
                         "advantaged group (non-Hispanic White, male, above poverty line)",
                         "disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                         ),
         group = factor(group, levels = c("disadvantaged group (non-White and/or Hispanic, female, below poverty line)",
                                          "advantaged group (non-Hispanic White, male, above poverty line)"))) |> 
  group_by(group, variable_grp) |> 
  summarize(mean_value = mean(abs(value)), .groups = "drop") |> 
  arrange(mean_value) |> 
  mutate(variable_grp = factor(variable_grp),
         variable_grp = fct_inorder(variable_grp)) |> 
  ggplot() +
  geom_bar(aes(x = variable_grp, y = mean_value, group = group, 
                 fill = group),  
             stat = "identity", position = "dodge") +
  labs(title = "Income",
       y = NULL,
       x = NULL,
       color = NULL) +
  theme(axis.text=element_text(size=9.5),
        axis.text.y = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
  scale_fill_manual(values = c("#cb6bce", "#240e31")) +
  coord_flip()

twoweek_lag <- twoweek_sex + twoweek_race + twoweek_income  + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom", legend.direction = "vertical") 

fig_s4 <- no_lag/twoweek_lag 
```

```{r}
#| label: fig-s4
#| fig-cap: "Global importance (mean |Shapley value|) for feature categories for no lag and 2-week lagged model by demographic group. The importance of each feature category for advantaged and disadvantaged demographic groups are displayed separately by color."
#| fig-height: 6
#| fig-width: 7
fig_s4
```


## Supplemental Tables

### Table S1: Demographic Contrasts for all Models

```{r}
#| label: tbl-fairness
#| tbl-cap: "This table reports the bayesian model contrasts for each demographic group comparison separately by model and contrast. The median value represents the difference in median area under the ROC curve (auROC) between the two groups. A negative value indicates lower auROC performance for the disadvantaged group (female, non-White and/or Hispanic, income below poverty line) compared to the advantaged group (male, non-Hispanic White, income above poverty line)."
#| code-fold: true

pp_sex <- pp_dem_contrast |> 
  filter(contrast == "female vs male") |> 
   mutate(ci = str_c("[", round(lower, 3), ", ", round(upper, 3), "]"),
         median = as.character(round(median, 3)),
         probability = as.character(round(probability, 3))) |> 
  select(lag, median, ci, probability) |> 
  rename(`Lag`= lag,
         Median = median,
         `Bayesian CI` = ci,
         Probability = probability)

pp_income <- pp_dem_contrast |> 
  filter(contrast == "below poverty vs above poverty") |> 
   mutate(ci = str_c("[", round(lower, 3), ", ", round(upper, 3), "]"),
         median = as.character(round(median, 3)),
         probability = as.character(round(probability, 3))) |> 
  select(lag, median, ci, probability) |> 
  rename(`Lag`= lag,
         `Median ` = median,
         `Bayesian CI ` = ci,
         `Probability ` = probability)

pp_race <- pp_dem_contrast |> 
  filter(contrast == "not white vs non-hispanic white") |> 
   mutate(ci = str_c("[", round(lower, 3), ", ", round(upper, 3), "]"),
         median = as.character(round(median, 3)),
         probability = as.character(round(probability, 3))) |> 
  select(lag, median, ci, probability) |> 
  rename(`Lag`= lag,
         `Median  ` = median,
         `Bayesian CI  ` = ci,
         `Probability   ` = probability)

pp_sex |> 
  full_join(pp_income, by = "Lag") |> 
  full_join(pp_race, by = "Lag") |> 
  kbl() |> 
  add_header_above(c(" " = 1, "Sex at birth" = 3, "Income" = 3, "Race/Ethnicity" = 3))
```


### Table S2: Brier scores for all models

```{r}
brier <- read_csv(here::here(path_models_lag, "brier_scores.csv"),
                  show_col_types = FALSE) |> 
  mutate(model = factor(model, 
                        levels = c("No lag", "1 day", "3 days",
                                   "1 week", "2 weeks"),
                        labels = c("No lag", "1-day", "3-day",
                                   "1-week", "2-week"))) |> 
  select(-beta)
  
brier |> 
  rename(Model = model, 
         `Raw (uncalibrated)` = raw,
         `Platt calibration` = logi) |> 
  knitr::kable(digits = 2) |> 
  kableExtra::kable_classic()
```



