---
title: "Make auROCs for Model Fairness Evaluations"
author: "Kendra Wyant"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  
  study: "lag"
  cv: "nested"
  window: "1week"
  model: "main" 
  lead: 336
  version: "v1"
editor_options: 
  chunk_output_type: console
---

### Set Up Environment

```{r}
study <- params$study
cv <- params$cv
window <- params$window
model <- params$model
lead <- params$lead 
version <- params$version
```

Function conflicts
```{r}
#| message: false
#| warning: false

# handle conflicts
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
```

Packages, functions, and paths
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_processed <- format_path(str_c("studydata/risk/data_processed/", study))
path_models <- format_path(str_c("studydata/risk/models/", study))
```


### Load model predictions

```{r}
preds_out <- read_rds(here::here(path_models, str_c("outer_preds_dem_6_x_5_", window, "_", 
                                                    lead, "_", version, "_", 
                                                    cv, "_", model, ".rds")))
```

### Load features and select demographics

```{r}
feat <- read_csv(here::here(path_processed, str_c("features_", lead, "lag_",
                                                  version, ".csv.xz")),
                 col_types = cols()) |> 
  select(label_num, subid, dttm_label, lapse, demo_age, demo_sex, demo_income, demo_race) |> 
  arrange(label_num) 
```

### Link features with predictions

```{r}
feat_preds <- feat |> 
  mutate(label = if_else(lapse == "yes", "Lapse", "No lapse"),
         id_obs = label_num) |>
  right_join(preds_out, by = c("id_obs", "label")) |> 
  select(-c(label_num, lapse)) |> 
  mutate(label = factor(label, levels = c("Lapse", "No lapse")))

# nrow(feat)
# nrow(preds_out) # should be nrow(feat) * 3
# nrow(feat_preds) # should be same as nrow(preds_out)
```

### Create dichotomous groups for auROC comparisons

Income, Gender, Race
```{r}
feat_preds <- feat_preds |> 
  mutate(demo_sex = if_else(demo_sex == "Female", "female", "male"),
         demo_race = if_else(demo_race == "White/Caucasian", "white", "not white"),
         demo_income = if_else(demo_income < 1560, "below poverty", "above poverty"))
```


### Calculate auROCs by group

This calculates auROC using test sets and probability corrected (prob_beta) probabilities.   

function
```{r}
get_auroc_by_dem <- function(df, var) {
  # df should be dataframe with demographic features and predictions
  # var should be character or factor variable in df
  
  var_tmp <- df |> 
      pull(get(var))
   
  var_levels <- 
     if (class(var_tmp) != "factor") {
       levels(as.factor(var_tmp))
     } else levels(var_tmp)
  
  # create empty tibble with outer split nums
  auroc <- tibble(outer_split_num = unique(df$outer_split_num)) |> 
    arrange(outer_split_num) 
    
  for (level in var_levels) {
     auroc_level <- df |>
       filter(get(var) == level) |> 
       nest(.by = outer_split_num, .key = "preds") |> 
       mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_beta, 
                                             truth = label))) |> 
       select(-preds) |> 
       unnest(auroc) |> 
       select(-c(.estimator, .metric)) |> 
       rename(!!level:=.estimate)
     
     auroc <- auroc |> 
       left_join(auroc_level, by = "outer_split_num")
  }
  
  return(auroc)
}
```


get auROCs 
```{r}
auroc_dem <- get_auroc_by_dem(df = feat_preds, var = "demo_sex") |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "demo_race"), by = "outer_split_num") |> 
  left_join(get_auroc_by_dem(df = feat_preds, var = "demo_income"), by = "outer_split_num")
```


Save auroc_dem
```{r}
auroc_dem |> 
  write_csv(here::here(path_models, str_c("test_auroc_dem_6_x_5_", window, "_", lead, 
                                          "_", version, "_", cv, ".csv")))  
```
