---
title: "Calculate auROC for first vs not-first lapses"
author: "Kendra Wyant"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

### Set Up Environment


Packages, functions, and paths
```{r}
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)
options(conflicts.policy = "depends.ok")

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_lag <- format_path("risk/data_processed/lag")
path_models <- format_path("risk/models/lag")
```


### Data

```{r}
preds_0 <- read_rds(here::here(path_models,
                             "outer_preds_6_x_5_1day_0_v3_nested_strat_lh.rds"))

labels_0 <- read_csv(here::here(path_lag, "labels_1day_0lag.csv"),
                   show_col_types = FALSE)

preds_24 <- read_rds(here::here(path_models,
                             "outer_preds_6_x_5_1day_24_v3_nested_strat_lh.rds"))

labels_24 <- read_csv(here::here(path_lag, "labels_1day_24lag.csv"),
                   show_col_types = FALSE)

preds_72 <- read_rds(here::here(path_models,
                             "outer_preds_6_x_5_1day_72_v3_nested_strat_lh.rds"))

labels_72 <- read_csv(here::here(path_lag, "labels_1day_72lag.csv"),
                   show_col_types = FALSE)

preds_168 <- read_rds(here::here(path_models,
                             "outer_preds_6_x_5_1day_168_v3_nested_strat_lh.rds"))

labels_168 <- read_csv(here::here(path_lag, "labels_1day_168lag.csv"),
                   show_col_types = FALSE)

preds_336 <- read_rds(here::here(path_models,
                             "outer_preds_6_x_5_1day_336_v3_nested_strat_lh.rds"))

labels_336 <- read_csv(here::here(path_lag, "labels_1day_336lag.csv"),
                   show_col_types = FALSE)
```

### Create grouping variable

- first_lapse = Yes if no lapses on study or all data up until and including first lapse
- first_lapse = No for all data after first lapse


```{r}
# get first lapses
(first_lapses_0 <- labels_0 |> 
  filter(lapse == "yes") |> 
  group_by(subid) |> 
  arrange(dttm_label) |> 
  slice_head(n = 1) |> 
  ungroup())

labels_0_full <- labels_0 |> 
  mutate(never_lapse = if_else(!subid %in% first_lapses_0$subid, "yes", "no"),
         first_lapse = "no")

for (the_subid in first_lapses_0$subid) {
  dttm <- first_lapses_0 |> 
    filter(subid == the_subid) |> 
    pull(dttm_label)
 
  # filter out full 24 hours of lapse
  
  row_ids <- which(
    labels_0_full$subid == the_subid &
    (labels_0_full$dttm_label <= dttm |
       (labels_0_full$dttm_label > dttm & labels_0_full$dttm_label < dttm + hours(24)))
  )

  labels_0_full[row_ids, "first_lapse"] <- "yes"
}

janitor::tabyl(labels_0_full$first_lapse)



labels_0_full |> 
  group_by(first_lapse) |> 
  summarise(n_subid = n_distinct(subid)) 

labels_0_full |> 
  filter(first_lapse == "yes" | never_lapse == "yes") |> nrow()

labels_0_full |> 
  filter(first_lapse == "no" | never_lapse == "yes") |> nrow()
```


Combine with preds
```{r}
labels_0_full <- labels_0_full |>
  mutate(label = if_else(lapse == "no", "No lapse", "Lapse"),
         id_obs = row_number())  |> 
  select(-lapse) |> 
  right_join(preds_0, by = c("id_obs", "label")) |> 
  mutate(label = factor(label, levels = c("Lapse", "No lapse")))
```



### Calculate auROCs comparing up first lapse to no lapse and not first lapse to no lapse

First lapses
```{r}
auroc_first_lapse_0 <- labels_0_full |> 
  filter(first_lapse == "yes" | never_lapse == "yes") |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) 
     
auroc_not_first_lapse_0 <- labels_0_full |> 
  filter(first_lapse == "no" | never_lapse == "yes") |>
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(not_first_lapse = .estimate) 

aurocs_0 <- auroc_first_lapse_0 |> 
  left_join(auroc_not_first_lapse_0, by = "outer_split_num") |> 
  arrange(outer_split_num) |> 
  glimpse()

aurocs_0 |> 
  write_csv(here::here(path_models, 
                       "test_auroc_6_x_5_1day_0_v3_nested_strat_lh_first_vs_later.csv"))
```


```{r}
round(median(aurocs_0$first_lapse), 2)
round(median(aurocs_0$not_first_lapse), 2)


# all data
labels_0_full |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) |> 
  pull(first_lapse) |> 
  median()
```


### Get lagged aurocs


```{r}
first_lapses_24 <- labels_24 |> 
  filter(lapse == "yes") |> 
  group_by(subid) |> 
  arrange(dttm_label) |> 
  slice_head(n = 1) |> 
  ungroup()

first_lapses_72 <- labels_72 |> 
  filter(lapse == "yes") |> 
  group_by(subid) |> 
  arrange(dttm_label) |> 
  slice_head(n = 1) |> 
  ungroup()

first_lapses_168 <- labels_168 |> 
  filter(lapse == "yes") |> 
  group_by(subid) |> 
  arrange(dttm_label) |> 
  slice_head(n = 1) |> 
  ungroup()

first_lapses_336 <- labels_336 |> 
  filter(lapse == "yes") |> 
  group_by(subid) |> 
  arrange(dttm_label) |> 
  slice_head(n = 1) |> 
  ungroup()

labels_24_full <- labels_24 |> 
  mutate(never_lapse = if_else(!subid %in% first_lapses_24$subid, "yes", "no"),
         first_lapse = "no")

labels_72_full <- labels_72 |> 
  mutate(never_lapse = if_else(!subid %in% first_lapses_72$subid, "yes", "no"),
         first_lapse = "no")

labels_168_full <- labels_168 |> 
  mutate(never_lapse = if_else(!subid %in% first_lapses_168$subid, "yes", "no"),
         first_lapse = "no")

labels_336_full <- labels_336 |> 
  mutate(never_lapse = if_else(!subid %in% first_lapses_336$subid, "yes", "no"),
         first_lapse = "no")

for (the_subid in first_lapses_24$subid) {
  dttm <- first_lapses_24 |> 
    filter(subid == the_subid) |> 
    pull(dttm_label)
 
  # filter out full 24 hours of lapse
  
  row_ids <- which(
    labels_24_full$subid == the_subid &
    (labels_24_full$dttm_label <= dttm |
       (labels_24_full$dttm_label > dttm & labels_24_full$dttm_label < dttm + hours(24)))
  )

  labels_24_full[row_ids, "first_lapse"] <- "yes"
}

for (the_subid in first_lapses_72$subid) {
  dttm <- first_lapses_72 |> 
    filter(subid == the_subid) |> 
    pull(dttm_label)
 
  # filter out full 24 hours of lapse
  
  row_ids <- which(
    labels_72_full$subid == the_subid &
    (labels_72_full$dttm_label <= dttm |
       (labels_72_full$dttm_label > dttm & labels_72_full$dttm_label < dttm + hours(24)))
  )

  labels_72_full[row_ids, "first_lapse"] <- "yes"
}

for (the_subid in first_lapses_168$subid) {
  dttm <- first_lapses_168 |> 
    filter(subid == the_subid) |> 
    pull(dttm_label)
 
  # filter out full 24 hours of lapse
  
  row_ids <- which(
    labels_168_full$subid == the_subid &
    (labels_168_full$dttm_label <= dttm |
       (labels_168_full$dttm_label > dttm & labels_168_full$dttm_label < dttm + hours(24)))
  )

  labels_168_full[row_ids, "first_lapse"] <- "yes"
}

for (the_subid in first_lapses_336$subid) {
  dttm <- first_lapses_336 |> 
    filter(subid == the_subid) |> 
    pull(dttm_label)
 
  # filter out full 24 hours of lapse
  
  row_ids <- which(
    labels_336_full$subid == the_subid &
    (labels_336_full$dttm_label <= dttm |
       (labels_336_full$dttm_label > dttm & labels_336_full$dttm_label < dttm + hours(24)))
  )

  labels_336_full[row_ids, "first_lapse"] <- "yes"
}
```


Combine with preds
```{r}
labels_24_full <- labels_24_full |>
  mutate(label = if_else(lapse == "no", "No lapse", "Lapse"),
         id_obs = row_number())  |> 
  select(-lapse) |> 
  right_join(preds_24, by = c("id_obs", "label")) |> 
  mutate(label = factor(label, levels = c("Lapse", "No lapse")))

labels_72_full <- labels_72_full |>
  mutate(label = if_else(lapse == "no", "No lapse", "Lapse"),
         id_obs = row_number())  |> 
  select(-lapse) |> 
  right_join(preds_72, by = c("id_obs", "label")) |> 
  mutate(label = factor(label, levels = c("Lapse", "No lapse")))

labels_168_full <- labels_168_full |>
  mutate(label = if_else(lapse == "no", "No lapse", "Lapse"),
         id_obs = row_number())  |> 
  select(-lapse) |> 
  right_join(preds_168, by = c("id_obs", "label")) |> 
  mutate(label = factor(label, levels = c("Lapse", "No lapse")))

labels_336_full <- labels_336_full |>
  mutate(label = if_else(lapse == "no", "No lapse", "Lapse"),
         id_obs = row_number())  |> 
  select(-lapse) |> 
  right_join(preds_336, by = c("id_obs", "label")) |> 
  mutate(label = factor(label, levels = c("Lapse", "No lapse")))
```

Calculate aurocs
```{r}
auroc_first_lapse_24 <- labels_24_full |> 
  filter(first_lapse == "yes" | never_lapse == "yes") |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) 
     
auroc_not_first_lapse_24 <- labels_24_full |> 
  filter(first_lapse == "no" | never_lapse == "yes") |>
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(not_first_lapse = .estimate) 

aurocs_24 <- auroc_first_lapse_24 |> 
  left_join(auroc_not_first_lapse_24, by = "outer_split_num") |> 
  arrange(outer_split_num) |> 
  glimpse()

aurocs_24 |> 
  write_csv(here::here(path_models, 
                       "test_auroc_6_x_5_1day_24_v3_nested_strat_lh_first_vs_later.csv"))

round(median(aurocs_24$first_lapse), 2)
round(median(aurocs_24$not_first_lapse), 2)


# all data
labels_24_full |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) |> 
  pull(first_lapse) |> 
  median()
```


```{r}
auroc_first_lapse_72 <- labels_72_full |> 
  filter(first_lapse == "yes" | never_lapse == "yes") |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) 
     
auroc_not_first_lapse_72 <- labels_72_full |> 
  filter(first_lapse == "no" | never_lapse == "yes") |>
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(not_first_lapse = .estimate) 

aurocs_72 <- auroc_first_lapse_72 |> 
  left_join(auroc_not_first_lapse_72, by = "outer_split_num") |> 
  arrange(outer_split_num) |> 
  glimpse()

aurocs_72 |> 
  write_csv(here::here(path_models, 
                       "test_auroc_6_x_5_1day_72_v3_nested_strat_lh_first_vs_later.csv"))

round(median(aurocs_72$first_lapse), 2)
round(median(aurocs_72$not_first_lapse), 2)


# all data
labels_72_full |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) |> 
  pull(first_lapse) |> 
  median()
```

```{r}
auroc_first_lapse_168 <- labels_168_full |> 
  filter(first_lapse == "yes" | never_lapse == "yes") |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) 
     
auroc_not_first_lapse_168 <- labels_168_full |> 
  filter(first_lapse == "no" | never_lapse == "yes") |>
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(not_first_lapse = .estimate) 

aurocs_168 <- auroc_first_lapse_168 |> 
  left_join(auroc_not_first_lapse_168, by = "outer_split_num") |> 
  arrange(outer_split_num) |> 
  glimpse()

aurocs_168 |> 
  write_csv(here::here(path_models, 
                       "test_auroc_6_x_5_1day_168_v3_nested_strat_lh_first_vs_later.csv"))

round(median(aurocs_168$first_lapse), 2)
round(median(aurocs_168$not_first_lapse), 2)


# all data
labels_168_full |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) |> 
  pull(first_lapse) |> 
  median()
```

```{r}
auroc_first_lapse_336 <- labels_336_full |> 
  filter(first_lapse == "yes" | never_lapse == "yes") |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) 
     
auroc_not_first_lapse_336 <- labels_336_full |> 
  filter(first_lapse == "no" | never_lapse == "yes") |>
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(not_first_lapse = .estimate) 

aurocs_336 <- auroc_first_lapse_336 |> 
  left_join(auroc_not_first_lapse_336, by = "outer_split_num") |> 
  arrange(outer_split_num) |> 
  glimpse()

aurocs_336 |> 
  write_csv(here::here(path_models, 
                       "test_auroc_6_x_5_1day_336_v3_nested_strat_lh_first_vs_later.csv"))

round(median(aurocs_336$first_lapse), 2)
round(median(aurocs_336$not_first_lapse), 2)


# all data
labels_336_full |> 
  nest(.by = outer_split_num, .key = "preds") |> 
  mutate(auroc = map(preds, \(preds) roc_auc(preds, prob_raw, 
                                             truth = label))) |> 
  select(-preds) |> 
  unnest(auroc) |> 
  select(-c(.estimator, .metric)) |> 
  rename(first_lapse = .estimate) |> 
  pull(first_lapse) |> 
  median()
```

