---
title: "callibration"
author: "Kendra wyant"
editor: visual
editor_options: 
  chunk_output_type: console
---

## Setup

```{r}
#| message: false

library(tidyverse)
library(patchwork)

theme_set(theme_classic())

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_models <- format_path(str_c("studydata/risk/models/lag"))
```

## Read in Probabilities (raw and calibrated)

```{r}
preds_0 <- read_rds(here::here(path_models, 
                               str_c(
                                 "outer_preds_3_x_10_1day_0_v3_nested_main.rds")))

preds_336 <- read_rds(here::here(path_models, 
                               str_c(
                                 "outer_preds_3_x_10_1day_336_v3_nested_main.rds")))
```

## View raw and logistic callibrated probabilities for no lag and 2-week lag

### No Lag

Raw probabilities
```{r}
preds_0 |> 
  ggplot(aes(x = prob_raw)) + 
   geom_histogram(bins = 100, fill = "white", col = "black") +
   # facet_wrap(~label, nrow = 2, scales = "free_y") +
   xlab("Pr(Lapse)") +
  theme(axis.text = element_text(size = rel(1.00)), 
        axis.title.x = element_text(size = rel(1.25)),
        strip.text = element_text(size = rel(1.75))) 
```

Logi calibrated
```{r}
preds_0 |> 
  ggplot(aes(x = prob_logi)) + 
   geom_histogram(bins = 100, fill = "white", col = "black") +
   # facet_wrap(~label, nrow = 2, scales = "free_y") +
   xlab("Pr(Lapse)") +
  theme(axis.text = element_text(size = rel(1.00)), 
        axis.title.x = element_text(size = rel(1.25)),
        strip.text = element_text(size = rel(1.75))) 
```

### 2-week Lag

Raw probabilities
```{r}
preds_336 |> 
  ggplot(aes(x = prob_raw)) + 
   geom_histogram(bins = 100, fill = "white", col = "black") +
   # facet_wrap(~label, nrow = 2, scales = "free_y") +
   xlab("Pr(Lapse)") +
  theme(axis.text = element_text(size = rel(1.00)), 
        axis.title.x = element_text(size = rel(1.25)),
        strip.text = element_text(size = rel(1.75))) 
```

Logi calibrated
```{r}
preds_336 |> 
  ggplot(aes(x = prob_logi)) + 
   geom_histogram(bins = 100, fill = "white", col = "black") +
   # facet_wrap(~label, nrow = 2, scales = "free_y") +
   xlab("Pr(Lapse)") +
  theme(axis.text = element_text(size = rel(1.00)), 
        axis.title.x = element_text(size = rel(1.25)),
        strip.text = element_text(size = rel(1.75))) 
```


828 logi probability scores > .75
```{r}
preds_336 |> 
  summarise(mean(prob_logi), min(prob_logi), max(prob_logi))

preds_336 |> 
  filter(prob_logi > .75) |> 
  nrow()
```

Compared to 17699 no lag scores over .75
```{r}
preds_0 |> 
  summarise(mean(prob_logi), min(prob_logi), max(prob_logi))

preds_0 |> 
  filter(prob_logi > .75) |> 
  nrow()
```


Similar pattern with raw scores - 2-week lag has only 25% of the predictions no lag has for probabilities over 75%
```{r}
preds_336 |> 
  filter(prob_raw > .75) |> 
  nrow()

preds_0 |> 
  filter(prob_raw > .75) |> 
  nrow()
```



## Calibration Plots

### Logi calibrated probabilities

```{r}
bin_width = 0.10

fig_cal_0 <- preds_0 |> 
  mutate(bins = cut(prob_logi, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "Lapse", 1, 0)) |> 
  group_by(bins)  |> 
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  geom_line() +
  geom_point() +
  labs(x = "Predicted Lapse Probability (bin mid-point)",
       y = "Observed Lapse Probability",
       title = "No Lag") +
  scale_x_continuous(breaks = seq(0, 1, bin_width),
                     limits = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, bin_width),
                     limits = c(0, 1)) 


fig_cal_336 <- preds_336 |> 
  mutate(bins = cut(prob_logi, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "Lapse", 1, 0)) |> 
  group_by(bins)  |> 
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  geom_line() +
  geom_point() +
  labs(x = "Predicted Lapse Probability (bin mid-point)",
       y = "Observed Lapse Probability",
       title = "2-week Lag") +
  scale_x_continuous(breaks = seq(0, 1, bin_width),
                     limits = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, bin_width),
                     limits = c(0, 1))
```

```{r}
fig_cal_0 + fig_cal_336
```

### No Calibration (raw probabilities)

```{r}
bin_width = 0.10

fig_cal_0 <- preds_0 |> 
  mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "Lapse", 1, 0)) |> 
  group_by(bins)  |> 
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  geom_line() +
  geom_point() +
  labs(x = "Predicted Lapse Probability (bin mid-point)",
       y = "Observed Lapse Probability",
       title = "No Lag") +
  scale_x_continuous(breaks = seq(0, 1, bin_width),
                     limits = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, bin_width),
                     limits = c(0, 1)) 


fig_cal_336 <- preds_336 |> 
  mutate(bins = cut(prob_raw, breaks = seq(0, 1, bin_width)), 
         lapse = if_else(label == "Lapse", 1, 0)) |> 
  group_by(bins)  |> 
  summarize(mean_lapse = mean(lapse),
            .groups = "drop") |>
  mutate(bins = as.numeric(bins),
         midpoints = bin_width/2 + bin_width * (bins - 1))  |> 
  ggplot(data = _, aes(x = midpoints, y = mean_lapse)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  geom_line() +
  geom_point() +
  labs(x = "Predicted Lapse Probability (bin mid-point)",
       y = "Observed Lapse Probability",
       title = "2-week Lag") +
  scale_x_continuous(breaks = seq(0, 1, bin_width),
                     limits = c(0, 1)) +
  scale_y_continuous(breaks = seq(0, 1, bin_width),
                     limits = c(0, 1))
```

```{r}
fig_cal_0 + fig_cal_336
```