---
title: "Supplemental Feature Plot Tibble"
author: "Kendra Wyant"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

### Set Up Environment


```{r, packages_workflow}
#| message: false
#| warning: false

# handle conflicts
options(conflicts.policy = "depends.ok")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_ml.R?raw=true")
tidymodels_conflictRules()
```


```{r, packages_script}
#| message: false
#| warning: false

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(tidymodels))
library(kableExtra, exclude = "group_rows")


theme_set(theme_classic()) 
```


```{r source_functions}
#| output: false

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
# CHTC support functions
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/chtc/static_files/fun_chtc.R?raw=true")
```

```{r, absolute_paths}
path_processed <- format_path(str_c("studydata/risk/data_processed/lag"))
path_models_lag <- format_path(str_c("studydata/risk/models/lag"))
path_chtc <- format_path(str_c("studydata/risk/chtc/lag/train_xgboost_1week_0lag_nested_1_x_10_3_x_10_v2_main/input"))

source(here::here(path_chtc, "training_controls.R"))
```


### Read in Data

Raw Shapley values - (testing with 0 lag models first)
```{r}
shap_0 <- read_rds(here::here(path_models_lag, 
                              "outer_shaps_1week_0_v1_nested_main.rds")) |> 
  glimpse()
```


Read in features
```{r}
feat_0 <- read_csv(here::here(path_processed, "features_0lag_v1.csv.xz"),
                   show_col_types = FALSE) |> 
  glimpse()
```

### Match raw feature to shaps

```{r}
feat_0 <- format_data(feat_0)  
rec <- recipe(y ~ ., data = feat_0) |>
    step_rm(subid) |>  
    step_impute_median(all_numeric_predictors()) |> 
    step_impute_mode(all_nominal_predictors()) |> 
    step_zv(all_predictors()) |> 
    step_dummy(all_factor_predictors()) |> 
    prep()

feat_0 <- bake(rec, new_data = NULL)
```


```{r}
feat_0 <- feat_0 |> 
  rename(id_obs = label_num) |> 
  select(-y)

feat_0 <- feat_0 |> 
  pivot_longer(cols = !starts_with("id"), names_to = "variable", values_to = "feature_score")
```

Update names of features to match SHAP
```{r}
clean_feature_names <- function(feat_name){
  new_name <- gsub(".l0", "", feat_name)
  new_name <- gsub("rratecount.count", "raw_count", new_name)
  new_name <- gsub("dratecount.count", "diff_count", new_name)
  new_name <- gsub("drecent_response", "diff_recent", new_name)
  new_name <- gsub("rrecent_response", "raw_recent", new_name)
  new_name <- gsub("dmin_response", "diff_min", new_name)
  new_name <- gsub("rmin_response", "raw_min", new_name)
  new_name <- gsub("dmax_response", "diff_max", new_name)
  new_name <- gsub("rmax_response", "raw_max", new_name)
  new_name <- gsub("dmedian_response", "diff_median", new_name)
  new_name <- gsub("rmedian_response", "raw_median", new_name)
  new_name <- gsub("label_", "", new_name)
  new_name <- gsub("demo_", "", new_name)
  new_name <- gsub("High.school.or.less", "high.school", new_name)
  new_name <- gsub("Some.college", "some.college", new_name)
  new_name <- gsub("Mon", "mon", new_name)
  new_name <- gsub("Tue", "tue", new_name)
  new_name <- gsub("Wed", "wed", new_name)
  new_name <- gsub("Thu", "thu", new_name)
  new_name <- gsub("Fri", "fri", new_name)
  new_name <- gsub("Sat", "sat", new_name)
  new_name <- gsub("Sun", "sun", new_name)
  new_name <- gsub("Never.Married", "never.married", new_name)
  new_name <- gsub("Never.Other", "never.other", new_name)
  new_name <- gsub("White.Caucasian", "caucasian", new_name)
  new_name <- gsub("Male", "male", new_name)
  new_name <- gsub("p12.raw_count.lapse", "lapse.p12.raw_count", new_name)
  new_name <- gsub("p24.raw_count.lapse", "lapse.p24.raw_count", new_name)
  new_name <- gsub("p48.raw_count.lapse", "lapse.p48.raw_count", new_name)
  new_name <- gsub("p72.raw_count.lapse", "lapse.p72.raw_count", new_name)
  new_name <- gsub("p168.raw_count.lapse", "lapse.p168.raw_count", new_name)
  new_name <- gsub("p12.diff_count.lapse", "lapse.p12.diff_count", new_name)
  new_name <- gsub("p24.diff_count.lapse", "lapse.p24.diff_count", new_name)
  new_name <- gsub("p48.diff_count.lapse", "lapse.p48.diff_count", new_name)
  new_name <- gsub("p72.diff_count.lapse", "lapse.p72.diff_count", new_name)
  new_name <- gsub("p168.diff_count.lapse", "lapse.p168.diff_count", new_name)
  new_name <- gsub("p12.raw_count.ema", "missing.p12.raw_count", new_name)
  new_name <- gsub("p24.raw_count.ema", "missing.p24.raw_count", new_name)
  new_name <- gsub("p48.raw_count.ema", "missing.p48.raw_count", new_name)
  new_name <- gsub("p72.raw_count.ema", "missing.p72.raw_count", new_name)
  new_name <- gsub("p168.raw_count.ema", "missing.p168.raw_count", new_name)
  new_name <- gsub("p12.diff_count.ema", "missing.p12.diff_count", new_name)
  new_name <- gsub("p24.diff_count.ema", "missing.p24.diff_count", new_name)
  new_name <- gsub("p48.diff_count.ema", "missing.p48.diff_count", new_name)
  new_name <- gsub("p72.diff_count.ema", "missing.p72.diff_count", new_name)
  new_name <- gsub("p168.diff_count.ema", "missing.p168.diff_count", new_name)
  return(new_name) 
}

feat_0 <- feat_0 |> 
  mutate(variable = fct_relabel(variable, clean_feature_names))
```

```{r}
nrow(shap_0)
nrow(feat_0)
```

Pull out one subject as example

```{r}
feat_test <- feat_0 |> 
  filter(id_obs == 1)

shap_test <- shap_0 |> 
  filter(id_obs == 1)

test <- feat_test |> 
  left_join(shap_test, by = c("id_obs", "variable"))
```

Not sure why there are two missing shap values for features for this observation?

For now combine will combine feature scores with shaps using right join

```{r}
shap_feat <- feat_0 |> 
  right_join(shap_0, by = c("id_obs", "variable")) |> 
  glimpse()
```

### Standardize feature scores

```{r}
shap_feat <- shap_feat |> 
  group_by(variable) |> 
  mutate(feature_score_z = (feature_score - mean(feature_score))/sd(feature_score))
```

### Group categories

Sum of shap values and mean of z-scored feature values

```{r}
shap_feat_grp <- shap_feat |>  
    mutate(variable_grp = if_else(str_detect(variable, "lapse."),
                           "past use (EMA item)",
                           variable),
           variable_grp = if_else(str_detect(variable_grp, "ema_2"),
                           "craving (EMA item)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "ema_3"),
                           "past risky situation (EMA item)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "ema_4"),
                           "past stressful event (EMA item)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "ema_5"),
                           "past pleasant event (EMA item)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "ema_6"),
                           "valence (EMA item)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "ema_7"),
                           "arousal (EMA item)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "ema_8"),
                           "future risky situation (EMA item)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "ema_9"),
                           "future stressful event (EMA item)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "ema_10"),
                           "future efficacy (EMA item)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "missing."),
                           "missing surveys (other)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "day"),
                           "lapse day (other)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "hour"),
                           "lapse hour (other)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "age"),
                           "age (demographic)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "sex"),
                           "sex (demographic)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "marital"),
                           "marital (demographic)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "race"),
                           "race (demographic)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "educ"),
                           "education (demographic)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "income"),
                           "income (demographic)",
                           variable_grp),
           variable_grp = if_else(str_detect(variable_grp, "employ"),
                           "employment (demographic)",
                           variable_grp)) |> 
    mutate(variable_grp = factor(variable_grp)) |> 
    group_by(id_obs, variable_grp) |>  # values are already averaged across repeats
    summarize(value = sum(value),
              feature_score_z_mean = mean(feature_score_z))
```


```{r}
 head(shap_feat_grp, n = 20)
```

### Save out tibble

```{r}
shap_feat_grp |> 
  write_rds(here::here(path_models_lag, "outer_shapsgrp_with_features_1week_0_v1_nested_main.rds")) 
```




