---
title: "Summarizes `r params$model` `r params$algorithms` models of inner loops for `r params$study` study"
author: "Kendra Wyant"
date: "March 24, 2024"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
params:
  study: "lag"
  version: "v2"
  algorithms: "xgboost"   # "all" or name of specific algorithm
  model: "main" # "main" or "baseline"
editor_options: 
  chunk_output_type: console
---



### Set Up Environment

```{r}
study <- params$study
version <- params$version
algorithms <- params$algorithms
model <- params$model
```

```{r}
#| message: false
#| warning: false

options(conflicts.policy = "depends.ok")
```

```{r}
#| message: false
#| warning: false

library(tidyverse)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
theme_set(theme_classic()) 
```


```{r}
#| message: false

source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_input <- format_path(str_c("studydata/risk/chtc/", study))
path_models <- format_path(str_c("studydata/risk/models/", study))
```


### Read results.csv files

Set up object for results

```{r}
results_all <- NULL
```

Get batch_names

```{r}
batch_names <- list.dirs(path_input, full.names = FALSE, recursive = FALSE)
  
if (algorithms == "all") {
  batch_names <- batch_names[str_detect(batch_names, "train") & 
                               str_detect(batch_names, version)]
  if (model == "main") {
    batch_names <- batch_names[!str_detect(batch_names, "baseline")]
  }
} else {
    batch_names <- batch_names[str_detect(batch_names, "train") & 
                               str_detect(batch_names, version) &
                               str_detect(batch_names, algorithms)]
  if (model == "main") {
    batch_names <- batch_names[!str_detect(batch_names, "baseline")]
  }
}  
 

# tmp filter out 336 lag
batch_names <- batch_names[!str_detect(batch_names, "336")]

batch_names
```

Loop over batch_names to read in files and perform checks

```{r}

for (batch_name in batch_names) {
  message("Processing Batch: ", batch_name)
  
  lag <- str_remove(str_split(batch_name, "_")[[1]][4], "lag")
  
  # read in configs
  configs <- read_csv(here::here(path_input, batch_name, "input", "configs.csv"), 
                      show_col_types = FALSE)
  (n_configs <- nrow(configs))
  
  # read in results
  results_batch <- read_csv(here::here(path_input, batch_name, "output", 
                                      "batch_results.csv"), 
                            show_col_types = FALSE) |> 
    mutate(lag = lag)
  (n_results_batch <- nrow(results_batch))
  
  # Check counts of results files
  if (!(n_configs == n_results_batch)) {
    stop(n_configs, " configs != ", n_results_batch, " results files!")
  } else {
    message(n_results_batch, " results files detected.  Correct!")
  }
  
  # Check col count
  if (!(ncol(results_batch) == 18)) {
    stop(ncol(results_batch), " columns != 18")
  } else {
    message(ncol(results_batch), " columns detected.  Correct!\n")
  }

  # Add batch to all metrics
  results_all <- results_all %>% 
    bind_rows(results_batch)
}
```

### Wrap up processing of raw metrics

Check for duplicates
```{r}
n_results <- nrow(results_all)

results_all <- results_all |> 
  distinct(split_num, outer_split_num, inner_split_num, algorithm, feature_set,
           hp1, hp2, hp3, resample, lag, .keep_all = TRUE)

if (nrow(results_all) != n_results) {
  stop(n_results - nrow(results_all), " duplicates detected")
}
```

Final checks across all batches

Check 300 performance metrics for all configurations
```{r}
results_all |> 
  group_by(algorithm, hp1, hp2, hp3, resample, lag) |> 
  count() |> 
  arrange(n) |> 
  head(n = 50)
```


### Median metrics across inner folds for model configurations

Inner loop performance of best config.
This median performance for each configuration over inner x outer folds (e.g., 300 folds for 1x10 inner and 3x10 outer). It is what we would get (essentially) if we just did simple k-fold but with LOTs of folds

```{r}
metrics_avg <- results_all %>% 
  group_by(algorithm, feature_set, hp1, hp2, hp3, resample, lag) %>% 
   summarize(across(c(accuracy, bal_accuracy, roc_auc, sens, spec, ppv, npv),
                    median),
             n_jobs = n(), .groups = "drop") %>% 
  relocate(n_jobs) %>% 
  arrange(desc(roc_auc)) |> 
  ungroup()

best_configs <- metrics_avg |> 
  group_by(lag) |> 
  slice(1) |> 
  print()
```

Performance metric plot across all inner folds

```{r}
for (i in 1:nrow(best_configs)) {
  best_config <- best_configs[i,]
  plot <- results_all |> 
    filter(algorithm == best_config$algorithm,
           feature_set == best_config$feature_set,
           hp1 == best_config$hp1,
           hp2 == best_config$hp2,
           hp3 == best_config$hp3,
           resample == best_config$resample) |> 
    ggplot(aes(x = roc_auc)) +
    geom_histogram(bins = 10) +
    labs(title = str_c(best_config$lag, " lag"))
  print(plot)
}

```
