---
title: "Model analyses"
author: "Kendra Wyant"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true 
    toc_depth: 4
format:
  html:
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

### Set Up Environment

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true"))
suppressPackageStartupMessages(library(tidyposterior))

path_models_lag <- format_path(str_c("studydata/risk/models/lag"))
path_shared <- format_path("studydata/risk/data_processed/shared")
path_processed <- format_path("studydata/risk/data_processed/lag")

options(knitr.kable.NA = '')
```


```{r}
test_metrics_0 <- read_csv(file.path(path_models_lag, 
                                        "test_metrics_1week_0_v1_nested.csv"), 
                              col_types = cols()) |> 
  filter(.metric == "roc_auc") |> 
  select(outer_split_num, "lag0" = .estimate)

test_metrics_24 <- read_csv(file.path(path_models_lag, 
                                       "test_metrics_1week_24_v1_nested.csv"),
                             col_types = cols()) |> 
  filter(.metric == "roc_auc") |> 
  select(outer_split_num, "lag24" = .estimate)

test_metrics_72 <- read_csv(file.path(path_models_lag, 
                                        "test_metrics_1week_72_v1_nested.csv"),
                              col_types = cols()) |> 
  filter(.metric == "roc_auc") |> 
  select(outer_split_num, "lag72" = .estimate)

test_metrics_168 <- read_csv(file.path(path_models_lag, 
                                        "test_metrics_1week_168_v1_nested.csv"), 
                              col_types = cols()) |> 
  filter(.metric == "roc_auc") |> 
  select(outer_split_num, "lag168" = .estimate)

test_metrics_336 <- read_csv(file.path(path_models_lag, 
                                       "test_metrics_1week_336_v1_nested.csv"),
                             col_types = cols()) |> 
  filter(.metric == "roc_auc") |> 
  select(outer_split_num, "lag336" = .estimate)

test_metrics_all <- test_metrics_0 |> 
  left_join(test_metrics_24, by = c("outer_split_num")) |> 
  left_join(test_metrics_72, by = c("outer_split_num")) |>
  left_join(test_metrics_168, by = c("outer_split_num")) |>
  left_join(test_metrics_336, by = c("outer_split_num")) |> 
  mutate(fold_num = rep(1:10, 3),
         repeat_num = c(rep(1, 10), rep(2, 10), rep(3, 10))) |> 
  select(-outer_split_num)
```


#### Model evaluation
```{r}
#| output: false


# Repeated CV (id = repeat, id2 = fold within repeat)
# with a common variance:  statistic ~ model + (model | id2/id)
set.seed(101)
pp <- test_metrics_all |> 
  rename(id = repeat_num,
         id2 = fold_num) |> 
  perf_mod(formula = statistic ~ model + (1 | id2/id),
         transform = tidyposterior::logit_trans,  # for skewed & bounded AUC
         iter = 2000, chains = 4, adapt_delta = .99, # defaults but may increase to fix divergence issues
         family = gaussian, 
)  
```

```{r}


test_metrics_all_perf <- test_metrics_all |>
  summarize(across(`lag0`:`lag336`, 
                   list(median = median,  
                        IQR = IQR)))

glimpse(test_metrics_all_perf)
```


The median auROC across the 30 test sets for our baseline model was high (median=`r sprintf("%1.3f", test_metrics_all_perf$lag0_median)`, IQR=`r sprintf("%1.3f", test_metrics_all_perf$lag0_IQR)`), consistent with our previous study. Performance across our lagged models was also high for the 24 hour lag (median=`r sprintf("%1.3f", test_metrics_all_perf$lag24_median)`, IQR=`r sprintf("%1.3f", test_metrics_all_perf$lag24_IQR)`), 72 hour lag (median=`r sprintf("%1.3f", test_metrics_all_perf$lag72_median)`, IQR=`r sprintf("%1.3f", test_metrics_all_perf$lag72_IQR)`), and 168 hour lag (median=`r sprintf("%1.3f", test_metrics_all_perf$lag168_median)`, IQR=`r sprintf("%1.3f", test_metrics_all_perf$lag168_IQR)`), and moderate for the 336 hour lag (median=`r sprintf("%1.3f", test_metrics_all_perf$lag336_median)`, IQR=`r sprintf("%1.3f", test_metrics_all_perf$lag336_IQR)`). 

```{r}


pp_tidy <- pp |> 
  tidy(seed = 123) 

q = c(.025, .5, .975)
test_metrics_all_pp_perf <- pp_tidy |> 
  group_by(model) |> 
  summarize(pp_median = quantile(posterior, probs = q[2]),
            pp_lower = quantile(posterior, probs = q[1]), 
            pp_upper = quantile(posterior, probs = q[3])) |> 
  mutate(model = factor(model, levels = c("lag0", "lag24", "lag72", "lag168", "lag336"),
                        labels = c("0 lag", "24 lag", "72 lag", "168 lag", "336 lag"))) |> 
  arrange(model)

test_metrics_all_pp_perf
```


Histograms of the full posterior probability distributions for auROC for each model are available in the supplement. The median auROCs from these posterior distributions were `r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "0 lag") |> pull(pp_median))` (baseline), `r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "24 lag") |> pull(pp_median))` (24 hour lag), `r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "72 lag") |> pull(pp_median))` (72 hour lag), `r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "168 lag") |> pull(pp_median))` (168 hour lag), and `r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "336 lag") |> pull(pp_median))` (336 hour lag). These values represent our best estimates for the magnitude of the auROC parameter for each model. The 95% Bayesian CI for the auROCs for these models were relatively narrow and did not contain 0.5: baseline[`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "0 lag") |> pull(pp_lower))`-`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "0 lag") |> pull(pp_upper))`], 24 hour lag[`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "24 lag") |> pull(pp_lower))`-`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "24 lag") |> pull(pp_upper))`], 72 hour lag[`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "72 lag") |> pull(pp_lower))`-`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "72 lag") |> pull(pp_upper))`], 168 hour lag[`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "168 lag") |> pull(pp_lower))`-`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "168 lag") |> pull(pp_upper))`], 336 hour lag[`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "336 lag") |> pull(pp_lower))`-`r sprintf("%1.3f", test_metrics_all_pp_perf |> filter(model == "336 lag") |> pull(pp_upper))`]. Figure 1 displays these median auROCs and 95% Bayesian CIs by model.

```{r}


test_metrics_all_pp_perf |>
  mutate(model = factor(model, levels = c("336 lag", "168 lag", "72 lag", "24 lag", "0 lag" ))) |> 
  ggplot() + 
  geom_point(aes(x = pp_median, y = model), color = "black") +
  geom_segment(mapping = aes(y = model, x = pp_lower, xend = pp_upper),
                data = test_metrics_all_pp_perf) +
  scale_x_continuous("Area Under ROC Curve", limits = c(.80, .925)) +
  ylab("Model") +
  theme_classic() +
  theme(legend.position = "none")
```



### Model Comparisons

#### Baseline Contrasts

```{r}


ci_baseline <- pp |>
  contrast_models(list("lag0", "lag0", "lag0", "lag0"), 
                  list("lag24", "lag72", "lag168", "lag336")) |> 
  summary(size = 0) |> 
  mutate(contrast = factor(contrast, 
                           levels = c("lag0 vs lag24", "lag0 vs lag72", "lag0 vs lag168", 
                                      "lag0 vs lag336"),
                           labels = c("0 vs. 24", "0 vs. 72", 
                                      "0 vs. 168", "0 vs. 336")))

ci_median_baseline <- pp |> 
  contrast_models(list("lag0", "lag0", "lag0", "lag0"), 
                  list("lag24", "lag72", "lag168", "lag336")) |>  
  group_by(contrast) |> 
  summarize(median = quantile(difference, .5)) |> 
  mutate(contrast = factor(contrast, 
                           levels = c("lag0 vs. lag24", "lag0 vs. lag72", "lag0 vs. lag168", 
                                      "lag0 vs. lag336"),
                           labels = c("0 vs. 24", "0 vs. 72", 
                                      "0 vs. 168", "0 vs. 336")))


ci_baseline <- ci_baseline |> 
  left_join(ci_median_baseline, by = c("contrast")) 

ci_baseline |> 
  select(contrast, probability, median, lower, upper) |> 
  arrange(contrast)
```

The median decrease in auROC for the baseline vs. 24 hour lag model was `r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 24") |> pull(median))` (95% CI=[`r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 24") |> pull(lower))`-`r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 24") |> pull(upper))`]), yielding a significant probability of `r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 24") |> pull(probability))` that the lagged model had worse performance. The median decrease in auROC for the baseline vs. 72 hour model was `r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 72") |> pull(median))` (95% CI=[`r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 72") |> pull(lower))`-`r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 72") |> pull(upper))`]), yielding a significant probability of `r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 72") |> pull(probability))` that the lagged model had worse performance. The median increase in auROC for the baseline vs. 168 hour lag model was `r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 168") |> pull(median))` (95% CI=[`r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 168") |> pull(lower))`-`r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 168") |> pull(upper))`]), yielding a significant probability of `r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 168") |> pull(probability))` that the lagged model had worse performance. The median increase in auROC for the baseline vs. 336 hour lag model was `r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 336") |> pull(median))` (95% CI=[`r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 336") |> pull(lower))`-`r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 336") |> pull(upper))`]), yielding a significant probability of `r sprintf("%1.3f", ci_baseline |> filter(contrast == "0 vs. 336") |> pull(probability))` that the lagged model had worse performance. 

#### Adjacent Contrasts

```{r}




ci_lag <- pp |>
  contrast_models(list("lag24", "lag72", "lag168"), 
                  list("lag72", "lag168", "lag336")) |> 
  summary(size = 0) |> 
  mutate(contrast = factor(contrast, 
                           levels = c("lag24 vs lag72", "lag72 vs lag168", 
                                      "lag168 vs lag336"),
                           labels = c("24 vs. 72", "72 vs. 168", "168 vs. 336")))

ci_median_lag <- pp |> 
  contrast_models(list("lag24", "lag72", "lag168"), 
                  list("lag72", "lag168", "lag336")) |>  
  group_by(contrast) |> 
  summarize(median = quantile(difference, .5)) |> 
  mutate(contrast = factor(contrast, 
                           levels = c("lag24 vs. lag72", "lag72 vs. lag168", 
                                      "lag168 vs. lag336"),
                           labels = c("24 vs. 72", "72 vs. 168", "168 vs. 336")))

ci_lag <- ci_lag |> 
  left_join(ci_median_lag, by = c("contrast")) 

ci_lag |> 
  select(contrast, probability, median, lower, upper) |> 
  arrange(contrast)
```


The median decrease in auROC for the 24 hour vs. 72 hour lag model was `r sprintf("%1.3f", ci_lag |> filter(contrast == "24 vs. 72") |> pull(median))` (95% CI=[`r sprintf("%1.3f", ci_lag |> filter(contrast == "24 vs. 72") |> pull(lower))`-`r sprintf("%1.3f", ci_lag |> filter(contrast == "24 vs. 72") |> pull(upper))`]), yielding a significant probability of `r sprintf("%1.3f", ci_lag |> filter(contrast == "24 vs. 72") |> pull(probability))` that the 72 hour lag model had worse performance than the 24 hour lag model. The median decrease in auROC for the 72 hour vs. 168 hour lag model was `r sprintf("%1.3f", ci_lag |> filter(contrast == "72 vs. 168") |> pull(median))` (95% CI=[`r sprintf("%1.3f", ci_lag |> filter(contrast == "72 vs. 168") |> pull(lower))`-`r sprintf("%1.3f", ci_lag |> filter(contrast == "72 vs. 168") |> pull(upper))`]), yielding a non-significant probability of `r sprintf("%1.3f", ci_lag |> filter(contrast == "72 vs. 168") |> pull(probability))` that the 168 hour lag model had worse performance than the 72 hour lag model. The median decrease in auROC for the 168 hour vs. 336 hour lag model was `r sprintf("%1.3f", ci_lag |> filter(contrast == "168 vs. 336") |> pull(median))` (95% CI=[`r sprintf("%1.3f", ci_lag |> filter(contrast == "168 vs. 336") |> pull(lower))`-`r sprintf("%1.3f", ci_lag |> filter(contrast == "168 vs. 336") |> pull(upper))`]), yielding a significant probability of `r sprintf("%1.3f", ci_lag |> filter(contrast == "168 vs. 336") |> pull(probability))` that the 336 hour lag model had worse performance than the 168 hour lag model. 
