---
title: "Preregistration: Lagged predictions of next week alcohol use" 
author: "Kendra Wyant"
date: "`r March 26, 2024"
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(source("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true"))

path_models_lag <- format_path(str_c("studydata/risk/models/lag"))
path_models_ema <- format_path(str_c("studydata/risk/models/ema"))
```


```{r}
#| label: ema-results

val_auc_week <- read_rds(file.path(path_models_ema, 
                                  "outer_metrics_1week_0_v5_nested_main.rds")) |> 
  select(outer_split_num, contains("auc_in"))

test_metrics_week <- read_csv(file.path(path_models_ema, 
                                        "test_metrics_1week_0_v5_nested.csv"), 
                              col_types = cols())

test_metrics_day <- read_csv(file.path(path_models_ema, 
                                       "test_metrics_1day_0_v5_nested.csv"),
                             col_types = cols())

test_metrics_hour <- read_csv(file.path(path_models_ema, 
                                        "test_metrics_1hour_0_v5_nested.csv"),
                              col_types = cols())
```

```{r}
#| label: inner-aucs

inner_lag <- read_csv(file.path(path_models_lag, 
                                 "inner_metrics_1week_0_v2_nested_main.csv"),
                       col_types = cols()) |> 
  mutate(lag = 0) |> 
  bind_rows(read_csv(file.path(path_models_lag, 
                                 "inner_metrics_1week_24_v2_nested_main.csv"),
                       col_types = cols()) |> 
  mutate(lag = 24)) |> 
  bind_rows(read_csv(file.path(path_models_lag, 
                                 "inner_metrics_1week_72_v2_nested_main.csv"),
                       col_types = cols()) |> 
  mutate(lag = 72)) |> 
  bind_rows(read_csv(file.path(path_models_lag, 
                                   "inner_metrics_1week_168_v2_nested_main.csv"),
                         col_types = cols()) |> 
  mutate(lag = 168)) |> 
  bind_rows(read_csv(file.path(path_models_lag, 
                                   "inner_metrics_1week_336_v2_nested_main.csv"),
                         col_types = cols()) |> 
  mutate(lag = 336)) |> 
  group_by(algorithm, feature_set, hp1, hp2, hp3, resample, lag) %>% 
    summarize(across(c(accuracy, bal_accuracy, roc_auc, sens, spec, ppv, npv),
                     median),
              n_jobs = n(), .groups = "drop") |> 
  group_by(lag) |> 
  arrange(desc(roc_auc)) |> 
  slice(1) |> 
  select(lag, algorithm, hp1, hp2, hp3, resample, roc_auc)
```

<!--Outstanding edits:
- Will need to change importance indices methods to use a model agnostic method for calculating shapley values.
- Create simulated data to demonstrate code. 

-->


### Prior Work

In a previous manuscript ([Wyant et al., in press](https://osf.io/preprints/psyarxiv/cgsf7)), we developed three separate models that provide hour-by-hour probabilities of future goal-inconsistent alcohol use (i.e., lapses) with increasing temporal precision: lapses in the next week, next day, and next hour. Model features were engineered from raw scores and longitudinal change in responses to 4X daily ecological momentary assessment (EMA). These features were derived to measure theoretically-implicated risk factors and contexts that have considerable support as predictors of lapses including past use, craving, past pleasant events, past and future risky situations, past and future stressful events, emotional valence and arousal, and self-efficacy. We demonstrated that it was possible to predict future alcohol lapses in the next week, day, and hour with high sensitivity and specificity (area under the receiver operating curves [auROCs] of `r sprintf("%1.2f", test_metrics_week |> filter(.metric == "roc_auc") |> pull(.estimate) |> median())`, `r sprintf("%1.2f", test_metrics_day |> filter(.metric == "roc_auc") |> pull(.estimate) |> median())`, and `r sprintf("%1.2f", test_metrics_hour |> filter(.metric == "roc_auc") |> pull(.estimate) |> median())` respectively). Due to class imbalance, however, positive predictive value (PPV) notably decreased as the prediction window became more narrow (PPVs of `r sprintf("%1.2f", test_metrics_week |> filter(.metric == "ppv") |> pull(.estimate) |> median())`, `r sprintf("%1.2f", test_metrics_day |> filter(.metric == "ppv") |> pull(.estimate) |> median())`, and `r sprintf("%1.2f", test_metrics_hour |> filter(.metric == "ppv") |> pull(.estimate) |> median())` for next week, day, and hour models respectively).

### Current Study

This study is an extension of our previous work. PPV, the proportion of positive predictions (i.e., lapses) that are true positives, is an important indicator of a model's clinical utility. A model that makes too many false positive predictions could result in loss of credibility of the model by the participant or patient, and could be costly if mobilizing resources to act on each positive prediction. PPV can be increased by increasing the decision threshold; however, increasing the decision threshold will also lower the model's sensitivity, or ability to correctly label positive cases as positive. With the high potential for harm if lapses are missed, decreasing sensitivity is not a viable solution for improving PPV. Therefore, of a prior three models, the week model with high sensitivity and moderate PPV is the most clinically useful model.    

Additionally, our previous models predicted imminent lapses. While, next hour and next day models are well-positioned to identify and recommend just-in-time interventions to address these immediate risks, the next week model may not have sufficient temporal specificity to recommend immediate patient action. We believe its clinical utility may improve if we shift this coarser window width into the future. This "time-lagged" model could provide patients with increased lead time to implement supports that might not be immediately available to them (e.g., schedule therapy appointment, request support from an AA sponsor).   

The current study proposes training a model to predict the probability of lapse at any point during a week window that is shifted one day, three days, one week, or two weeks into the future. We will use our imminent lapse (i.e., no lag) week model as a baseline comparison and determine whether performance significantly degrades at each time shift into the future.     


### Methods
#### Participants
We recruited participants in early recovery (1-8 weeks of abstinence) from moderate to severe alcohol use disorder in Madison, Wisconsin, US for a three month longitudinal study. We used data from all participants included in our previous study (N = 151; see [Wyant et al., in press](https://osf.io/preprints/psyarxiv/cgsf7) for enrollment, disposition, and demographic information).   

#### Lapse Labels
We predicted lapses occurring in one week windows. Predictions were updated hourly. Therefore our models provided hour-by-hour predictions of future probability of an alcohol lapse within a one week window shifted at various time points (i.e., lags) into the future (no lag and one day, three day, one week, or two week lagged predictions). The first prediction for each participant occurs 24 hours + lag time (in hours) after their study start date. For example, the first prediction for the one day lagged prediction model would be 48 hours after participants' start date.

A prediction window was labeled *lapse* if the start date/hour of any drinking episode fell within that window.  A window was labeled *no lapse* if no alcohol use occurred within that window +/- 6 hours.  If no alcohol use occurred within the window but did occur within 6 hours of the start or end of the window, the window was excluded. We had previously used a more conservative fence around lapse exclusions (24 hours vs. 6 hours), however, after re-evaluation we deemed this to be too conservative. Inner loop auROCs of our initial no lag models suggest similar performance to our previous 24 hour fence (`r sprintf("%1.2f", inner_lag |> filter(lag == 0) |> pull(roc_auc))` vs. `r sprintf("%1.2f", val_auc_week |> pull(roc_auc_in) |> median())` auROC respectively).


#### Feature Engineering
Features were calculated using only data collected before the start of each prediction window. For our no lag models this included all data prior to the hour of the start of the prediction window. For our lagged models, the last EMA data used for feature engineering were collected up to one day, three days, one week, or two weeks prior to the start of the prediction window.   

Following our previous work, features were derived from three data sources:    

- *Demographics*: We created quantitative features for age and personal income, and dummy-coded features for sex, race/ethnicity, marital status, education, and employment.    

- *Day and time of prediction window onset*: We created dummy-coded features to indicate time of day and day of week that the prediction window began.    

- *Previous EMA responses*: We created raw EMA and change features for varying scoring epochs before the start of the prediction window for all EMA items (i.e., 12, 24, 48, 72, and 168 hours). Raw features included min, max, and median scores for each EMA question across all EMAs in each epoch for that participant. We calculated change features by subtracting the participant's mean score for each EMA question (using all EMAs collected before the start of the prediction window) from the associated raw feature. We also created raw and change features based on the most recent response for each EMA question and raw and change rate features from previously reported lapses.

#### Modeling Decisions
Our primary performance metric for model selection and evaluation will be auROC.   

We will consider four candidate statistical algorithms including elastic net, XGBoost, regularized discriminant analysis (rda), and single layer neural networks. This will ensure we are considering algorithms that differ in terms of bias-variance tradeoff (e.g., parametric vs non-parametric) to select the model configuration that most closely represents the true data generating process.   

We will use participant-grouped, nested cross-validation for model training, selection, and evaluation with auROC. We will use 1 repeat of 10-fold cross-validation for the inner loops (i.e., *validation* sets) and 3 repeats of 10-fold cross-validation for the outer loop (i.e., *test* sets).    

Best model configurations will be selected using median auROC across the 10 validation sets.  Final performance evaluation of those best model configurations used median auROC across the 30 test sets.     


### Progress at Time of Preregistration
Labels and features for all five models (no lag and one day, three day, one week, or two week lagged predictions) are created. 

The inner nested k-fold cross-validation loops have been run for XGBoost for all five models. These folds serve as validation sets for selecting the best performing model configuration (i.e., the combination of algorithm, hyperparameter values, and resampling method).   

- *algorithm* - In our prior study, preliminary testing suggested XGBoost was the best model configuration from a handful of algorithms, including Random Forest, K-Nearest Neighbors, and Elastic Net. Thus, we began with training XGBoost model configurations. Next, we will train our remaining candidate statistical algorithms (elastic net, rda, and neural networks).    

- *Hyperparameter values* - Candidate XGBoost model configurations differed across sensible values for the XGBoost hyperparameters mtry, tree depth, and learning rate. All configurations used 500 trees with early stopping to prevent over-fitting. All other hyperparameters were set to tidymodels package defaults. Our elastic net model configurations will differ across sensible values for the hyperparameters lambda and alpha. Our rda model configurations will differ across sensible values for the hyperparameters gamma and delta. Our neural network models will differ across sensible values for number of hidden units, and regularization values for penalty and dropout. We will also use two activation functions (relu and softmax). All neural network configurations will use the Adam optimizer for tuning learning rate. 

- *Resampling method* - Candidate model configurations also differed on outcome resampling method (i.e., up-sampling and down-sampling of the outcome using majority/no lapse to minority/lapse ratios ranging from 1:1 to 3:1).   

**No performance metrics have been calculated in the 30 outer folds (i.e., our test sets for model evaluation) and no planned data analyses have been done.**     


The best model configurations for each model selected from the inner loop validation sets are detailed below.  

```{r}
inner_lag |> 
  kableExtra::kbl(digits = 3) |> 
  kableExtra::kable_classic()
```


### Planned Data Analyses

#### Model Evaluation

We will use a Bayesian hierarchical generalized linear model to estimate the posterior probability distributions and 95% Bayesian confidence intervals (CIs) for auROC for the five best models (no lag and 1 day, 3 day, 1 week, and 2 week lagged predictions). The median auROCs from these distributions will represent our best estimate for the magnitude of the auROC parameter for each model. We will also examine whether the 95% Bayesian CI for the auROCs for these models contain 0.5 (chance performance). If the confidence intervals do not contain .5, this suggests our model is capturing signal in the data. 

#### Model Comparisons

We will perform four Bayesian model comparisons to determine the probability that the lagged models systematically degrade as predictions are made further into the future (i.e., longer lag time). Our four model comparisons will consist of the following repeated contrasts:    

1. No lag vs. 1 day lag.
2. 1 day lag vs. 3 day lag.
3. 3 day lag vs. 1 week lag.
4. 1 week lag vs. 2 week lag.

To determine the probability that the models' performances differed systematically from each other, we will regress the auROCs (logit transformed) from the 30 test sets for each model as a function of contrast (No lag vs. 1 day, 1 day vs. 3 day, 3 day vs. 1 week, 1 week vs. 2 week). We will set two random intercepts: one for the repeat, and another for the fold within repeat (folds are nested within repeats for 3x10-fold cross-validation). We will also report 95% (equal-tailed) Bayesian CIs for the differences in performance associated with the Bayesian comparisons. 

#### Feature Importance Indices

In our previous study we used a XGBoost-specific method for calculating feature importance indices. A benefit of using an ensemble method like XGBoost is that it requires low computational cost to get estimates of feature importance from a trained model. Therefore, if XGBoost is our best candidate statistical algorithm, we will use the SHAPforxgboost package designed for XGBoost models. The SHAPforxgboost package provides Shapley values in log-odds units for binary classification models.

If our best model configuration uses an algorithm other than XGBoost we will use a model-agnostic, more computationally expensive, method for calculating Shapley values. Specifically, we will **add explanation of method.**

Using one of the two methods above, we will calculate Shapley values from the 30 test sets to provide an explanation of the importance of categories of features across our five models. We will average the three Shapley values for each observation for each feature (i.e., across the three repeats) to increase their stability. An inherent property of Shapley values is their additivity, allowing us to combine features into feature categories. We will create separate feature categories for each of the nine EMA questions, the rates of past alcohol use and missing surveys, the time of day and day of the week of the start of the prediction window, and the seven demographic variables included in the models.

To calculate the local (i.e., for each observation) importance for each category of features, we will add Shapley values across all features in a category, separately for each observation.  To calculate global importance for each feature category, we will average the absolute value of the Shapley values of all features in the category across all observations. These local and global importance scores based on Shapley values allow us to contextualize relative feature importance.    

### Data Simulation Example

Below we simulate 30 folds of random data to demonstrate the code and feasibility of our planned analyses. 












